### 2025-04-02 paper/main.tex
Revamp Discussion

\documentclass{WileyMSP-template}

\pagestyle{plain}

\title{\Large\textbf{Advection versus diffusion in brain ventricular transport}}

\author[1,*]{Halvor Herlyng}
\author[1]{Ada J.~Ellingsrud}
\author[1]{Miroslav Kuchta} 
\author[2]{Inyoung Jeong}
\author[1,3]{Marie E.~Rognes}
\author[2,*]{Nathalie Jurisch-Yaksi}
\affil[1]{Department of Numerical Analysis and Scientific Computing, Simula Research Laboratory, Oslo, Norway} % Kristian Augusts gate 23, 0164
\affil[2]{Department of Clinical and Molecular Medicine, Norwegian University of Science and Technology, Trondheim, Norway} % Erling Skjalgsons gate 1, 7491
\affil[3]{K.~G.~Jebsen Center for Brain Fluid Research, Oslo, Norway}
\affil[*]{\href{mailto:hherlyng@simula.no}{hherlyng@simula.no}, \href{mailto:nathalie.jurisch-yaksi@ntnu.no}{nathalie.jurisch-yaksi@ntnu.no}}
\date{\vspace{-2em}} % Remove space between affiliations and abstract

\begin{document}

\maketitle

\begin{abstract}
Cerebrospinal fluid (CSF) is integral to brain function, as the fluid
provides mechanical support for the brain and the flow of CSF helps
distribute nutrients, neurotransmitters and metabolites throughout the
central nervous system. CSF flow is partly driven by beating of motile
cilia, hair-like structures that line the ependymal cell layer of the
walls of the brain ventricles. Despite the physiological importance of
CSF flow, the underlying mechanisms of cilia-mediated flow and
transport in the brain ventricles remain yet to be comprehensively
resolved. This study analyzes and evaluates the role of motile cilia
in brain intraventricular CSF flow and transport. The geometry of the
zebrafish larval brain ventricles, for which detailed knowledge of
cilia properties and CSF motion velocities and frequencies exists, was
used to develop finite element methods for modelling flow and
transport. The computational model is validated by in-vivo experiments
that monitor transport of a photoconvertible protein secreted in the
brain ventricles. Our results show that while cilia contribute to
advection of large particles, diffusion plays a significant role in
the transport of small solutes. We also demonstrate how cilia location
and the geometry of the ventricular system impact solute
distribution. Altogether, this work presents a computational framework
that can be applied to other ventricular systems, together with new
concepts of how molecules are transported within the brain and its
ventricles.

\end{abstract}

\noindent\small\hspace{.8725cm}\textbf{Keywords} cerebrospinal fluid, brain ventricles, cilia, transport, finite elements

\section{Introduction}

% I1: What is the problem - from big picture to our precise topic - why is it interesting?
Brain development and function depend on numerous regulatory
processes, including the flow of cerebrospinal fluid
(CSF)~\cite{fame2020emergence}. Produced mostly by the choroid
plexus and partly by brain interstitial fluid,
CSF is a clear liquid that flows through the ventricular system and subarachnoid
space~\cite{Brodal2016, Damkier2013}. By transporting growth factors, nutrients, neurotransmitters
and metabolites, as well as removing waste products, CSF regulates
brain development and homeostasis~\cite{Ringers2020Role, fame2020emergence, del2010ependymal}.
Disruption of the CSF circulation is detrimental to brain health and disturbs
cognitive and motor functions~\cite{Johanson2008}.
Pathological CSF flow may also lead to
disorders such as hydrocephalus, where accumulation of
CSF often requires emergency surgical
treatment~\cite{duy2022rethinking, kahle2024paediatric, wallmeier2022role},
but for which there is little agreement on optimal treatments~\cite{flannery2014pediatric}.

% I2: What do we know about the problem and the topic?
The flow of CSF in the ventricular system is driven by multiple
physiological processes, including via the beating motion of motile
cilia, CSF secretion by the choroid plexus, and pulsatile pressure
gradients generated by the cardiac cycle, respiration, neural
activity, and body movements~\cite{Brodal2016, del2010ependymal,
  Ringers2020Role, Vinje2019RespiratoryMeasurements,
  Kurtcuoglu2005ComputationalSystem, Olstad2019CiliaryDevelopment,
  macaulay2022cerebrospinal, mestre2018flow}.  Cilia are hair-like
structures found in a vast diversity of biological organisms and in
several human organs, including the brain~\cite{DGama2025MotileBrain,
  Fulford1986Muco-ciliaryLung, Pacherres2022CiliaryProduction,
  Jahn1972LocomotionProtozoa, Brennen1977FluidFlagella,
  Ringers2020Role, spassky2013motile, Pellicciotta2020Cilia,
  mitchell2007evolution, Tsukita2012CoordinatedFeet,
  BLAKE1974MechanicsMotion,
  Reiten2017Motile-Cilia-MediatedComputations,
  Thouvenin2020OriginCanal}.  The ventricular surface of the brain,
known as the ependyma, is lined by motile cilia that beat in a
coordinated pattern at frequencies of 10--40 Hz to mediate fluid
flow~\cite{mitchell2007evolution, Ringers2020Role, spassky2013motile,
  Ringers2023NovelEpithelia, roth2025structure}.  Given their
microscopic size, ranging from 5 to 15 microns, cilia contribute
primarily to flow near the ependyma in the human
brain~\cite{Siyahhan2014FlowVentricles, Olstad2019CiliaryDevelopment,
  Ringers2020Role, Faubel2016Cilia-basedVentricles}. In contrast,
cardiac, respiratory and delta waves at lower frequencies induce
pulsatile flow at the larger scales in the brain ventricular system,
subarachnoid space, and subarachnoid perivascular
spaces~\cite{Vinje2019RespiratoryMeasurements, eide2024functional,
  Causemann2025, fultz2019coupled, mestre2018flow,
  daversin2020mechanisms}.

% I3: What do we not know about the problem and what do we want to
% know? What are the precise research questions we want to address?
% Refer to specific studies and what specifically we do not know.
To date, it remains poorly understood how motile cilia contribute to 
the overall fluid flow and solute transport within the brain.
Experimental approaches face technical challenges: the microscopic size and relatively 
high beating frequencies of the cilia~\cite{ocallaghan2012analysis, mahuzier2018ependymal,
Jeong2022MeasurementTelencephalon, Olstad2019CiliaryDevelopment, spassky2013motile, DGama2021Diversity},
as compared to other fluid dynamic processes in the brain that are slower and span larger
spatial scales~\cite{Vinje2019RespiratoryMeasurements, eide2024functional, fultz2019coupled, mestre2018flow},
make cilia difficult to monitor in vivo in humans.
Until now, most studies have analyzed one physical mechanism at a time, especially in mammals.
For instance, studies using brain explants have convincingly shown
that cilia generate complex flow patterns at the surface of the brain
ventricles~\cite{Faubel2016Cilia-basedVentricles, Sawamoto2006NewBrain, ibanez2003beat};
however, they did not study the role of other contributions to CSF
flow and the accompanying transport, such as pressure
gradients associated with respiration and cardiac pulsations. Moreover, while MRI technology
can measure CSF motion in humans~\cite{feinberg1987human, battal2011cerebrospinal,
yamada2015current}, it cannot detect the microscopic
movements of cilia to understand their impact in the large human
brain~\cite{Siyahhan2014FlowVentricles}. 
These technical limitations of studying cilia-mediated flow and transport make 
mathematical modeling and computational investigations an attractive complement to experimental
studies. Previous simulation studies have focused on modeling the interaction
between single or several cilia and surrounding
fluids~\cite{Guo2020SimulatingGeometries, Ruvalcaba2021NumericalTree, Smith2009MathematicalFluids,
Cui2019NumericalMethod, Cui2022AFlow}.
Others considered cilia-mediated flow in the brain~\cite{Siyahhan2014FlowVentricles,
Yoshida2022EffectVentricles, Salman2022ComputationalEmbryo, Thouvenin2020OriginCanal}, but also 
other applications such as ciliary transport of mucous and particles in
airways~\cite{Fulford1986Muco-ciliaryLung, Ramirez-SanJuan2020Multi-scaleArrays, roth2025structure},
propulsion and locomotion of ciliated organisms~\cite{BLAKE1974MechanicsMotion,
Jahn1972LocomotionProtozoa, Brennen1977FluidFlagella}, and
oxygen regulation in coral reefs~\cite{Pacherres2022CiliaryProduction}.
These simulation studies have provided valuable insights on how cilia mediate complex
biological flows, and how cilia may contribute to tissue physiology.
However, we still lack a comprehensive understanding of how
cilia contribute to CSF motion and solute transport within the brain.
This is of particular importance since disorders such as hydrocephalus are observed
in animal models and humans with deficient motile cilia~\cite{Afzelius2004CiliaRelatedDiseases,
Olstad2019CiliaryDevelopment, Mitchison2017Motile, Ibanez-Tallon2002LossHydrocephalus},
indicating that the cilia play a crucial role for healthy brain function.

% I4: How do we address the problem and what do we find out?
In this study, we combined computational and experimental approaches to investigate
the role of motile cilia in intraventricular CSF flow and transport.
We model the flow and transport by partial differential equations and solve
them numerically with finite element methods to compute flow patterns and
solute concentration dynamics. A geometry representing the zebrafish larval brain ventricles
was used, for which we have detailed knowledge of cilia properties and CSF motion velocities
and frequencies~\cite{Olstad2019CiliaryDevelopment}. 
Notably, this early developmental stage allows us to simplify our model
by omitting CSF secretion, which starts at later stages~\cite{Jeong2024TheZebrafish},
while still being well-conserved with mammals~\cite{DGama2025MotileBrain,
Olstad2019CiliaryDevelopment, Ringers2020Role, DGama2021Diversity, jurisch2020radial}. 
The computational model is validated against experimental data by monitoring in vivo
the dynamics of a photoconvertible fluorescent protein Dendra2
expressed within the brain ventricles.
Our results show that diffusion plays a major role in the transport of small solutes,
whereas cilia are paramount for the movement of larger particles.
We also found that the location of cilia, as well as changes to the ventricular geometry,
largely impact solute distribution.
In summary, our work presents a computational
framework that can be applied to other
ventricular systems in the future, together with new concepts
of how cilia distribute molecules and
particles within the brain ventricles.

\section{Results}

\subsection{Motile cilia induce flow compartmentalization within the brain ventricles}

% Why? (Interesting point #1. However, we do not know aspect #2. 
% How? (We therefore ...)
% What ...

In vivo particle tracking reveals characteristic CSF flow patterns in
zebrafish larval brain ventricles
(\Crefpanels{fig:fig1}{a--c})~\cite{Olstad2019CiliaryDevelopment}.
We attempt to accurately represent these flow patterns with
computational fluid dynamics driven by cardiac pulsations and motile
cilia lining the surface of the brain ventricles.

Assuming that viscous forces are dominating the intraventricular CSF flow,
we model the flow by the Stokes equations.
The net forces of motile cilia acting on the CSF are modeled as a steady,
bidirectional tangential traction acting on parts of the ventricular wall
(\Crefpanels{fig:fig1}{d}), and we assume free velocity slip at the walls.
Simulations show that this traction alone
leads to a partial compartmentalization of the ventricular system,
with large-scale vortex structures in the anterior, middle and
posterior ventricles (\Crefpanels{fig:fig1}{e}). The highest flow
speeds ($26.7$ \textmu m/s) occur in the vicinity of the dorsal cilia
in the middle ventricle. To model the cardiac pulsations, we impose a
pulsatile pressure difference between the anterior and posterior
boundaries (\Crefpanels{fig:fig1}{d, f}). The resulting flow
is pulsatile and directional (\Crefpanels{fig:fig1}{g}), and
in the absence of the cilia, the vortex structures vanish. The
velocity magnitude peaks at $4.8$ \textmu m/s. CSF initially flows
rostrocaudally before changing direction to caudorostral in the middle
of the cardiac cycle.

When combining cilia motion and cardiac pulsations, we recognize flow features
persisting throughout the cardiac cycle, now with vortex structures in
the anterior and middle ventricles and directional flow in the
posterior ventricle (\Crefpanels{fig:fig1}{h--j}, \fixme{Supplementary
  video 1}). These characteristic patterns suggest that the flow in
the anterior and middle ventricles is primarily driven by the motile
cilia, while the cardiac pressure pulsations dominate within the
posterior ventricle and in the ducts connecting the ventricular
compartments. The CSF flow speed peaks at the time of peak caudal
cardiac pulsatile flow, when both the tangential traction induced by
the cilia and the pulsatile pressure-driven flow are aligned with the
rostrocaudal axis, reaching up to 27.9 \textmu m/s dorsally in the
posterior part of the middle ventricle. The flow Reynolds number is
$\mathrm{Re} \approx 0.004$, justifying the assumption of Stokes
flow. In summary, the computational CSF flow model reproduces the flow
features observed in vivo~\cite{Olstad2019CiliaryDevelopment}.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{graphics/figure1.png}
    \caption{
    \textbf{a.} Schematic illustration of the two cerebrospinal fluid (CSF) flow components modeled,
    motile cilia and cardiac pulsations, and their contributions to the flow patterns.
    \textbf{b, c.} Particle tracking data visualizing CSF flow directionality (\textbf{b}) and
    velocity magnitudes (\textbf{c}) in the middle ventricle.
    \textbf{d.} Computational mesh with marked facets indicating regions of motile cilia (magenta) and
    anterior/posterior facets (blue). Dorsal (D), ventral (V), anterior (A), posterior (P),
    left (L) and right (R) coordinates indicated.
    \textbf{e.} Streamlines of the (steady-state) CSF flow field simulated with the
    cilia-driven/no-cardiac flow model.
    \textbf{f.} Illustration of the cardiac-related normal pressure forcing imposed on the
    anterior/posterior boundaries. Time instants of streamline plots indicated.
    \textbf{g.} Streamlines of the CSF flow field simulated with the cardiac-induced/no-cilia flow
    model at a time 25\% into the cardiac cycle. %t=0.112613 s, period=0.45045 s
    \textbf{h--j.} Streamlines simulated with the baseline flow model (cilia+cardiac)
    at a time 25\% (\textbf{h}), 50\% (\textbf{i}) and 75\% (\textbf{j})
    into the cardiac cycle. %t=0.112613 s %t=0.225225 s %t=0.315315 s
    Scale bars 50 \textmu m.
    }
    \label{fig:fig1}
\end{figure}

\subsection{Ventricular solute transport is advection-dominated for larger molecules}

Proteins, neurotransmitters and other molecular substances
intrinsically diffuse within the brain ventricles, but are also
advected by the flow of CSF. A natural question is which of these
mechanisms that dominate for any given type of solute. To address this
question, we simulate the distribution and evolution of three
different diffusing solutes injected within a region of interest (ROI 1) in the
dorsal region of the middle ventricle (\Crefpanels{fig:fig2}{a}),
subject to the CSF flow induced by motile cilia and cardiac pressure
pulsations. In particular, we consider three different molecular
diffusion coefficients, corresponding to a large extracellular vesicle (EV, $D_1 = 2.17
\times 10^{-12}$ m$^2$/s), a 90 kDa Starmaker-Green Fluorescent Protein
(STM-GFP, $D_2 = 5.75 \times 10^{-11}$ m$^2$/s),
and a 26 kDa photoconvertible protein
Dendra2 ($D_3 = 1.15 \times 10^{-10}$ m$^2$/s).

Interestingly, the distribution patterns differ between the
solutes. For the smallest diffusion coefficient, representing
extracellular vesicles, the advective transport by the CSF dominates diffusion: the
CSF flow structures carry over into the concentration field,
indicating that transport along the streamlines is more rapid than the
diffusion across the streamlines (\Crefpanels{fig:fig2}{b}). The
balance between diffusion and advection shifts with increasing
diffusion coefficient as the solute spreads more uniformly throughout
the ventricular geometry (\Crefpanels{fig:fig2}{b--d}). Notably, the
smaller solutes with higher diffusion coefficients spread more quickly
to distal regions. Computing Péclet numbers, we find that
$\mathrm{Pe}_1=664$, $\mathrm{Pe}_2=25$ and $\mathrm{Pe}_3=13$ for
$D_1$, $D_2$ and $D_3$, respectively. With all $\mathrm{Pe} > 1$, this
indicates that advection is the dominant transport mechanism on a
global scale.

\begin{figure}%[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure2_simulation_results_different_coeffs.png}
    \caption{Simulated distribution and evolution after photoconversion of
    extracellular vesicles (EVs, $D_1$), STM-GFP ($D_2$) and Dendra2 ($D_3$). 
    \textbf{a.} Geometry-centered clip of the ventricles mesh showing the locations of the 
    regions of interest (ROIs).
    \textbf{b--d.} Slices of the concentration field simulated
    with diffusion coefficients $D_1$, $D_2$ and $D_3$ for nine time instants covering the
    time range from 90 seconds to 811 seconds, corresponding to 200 to 1800 cardiac cycles,
    for even intervals of 200 cycles.
    Slices are taken in the $xz$-plane ($y=0.140$ mm, the center of the geometry).
    \textbf{e--g.} Mean concentration $\cbar$ in each ROI
    as function of time for diffusion coefficients $D_1$, $D_2$ and $D_3$. 
    \textbf{h.} The time $\hat{t}$ before the mean concentration $\cbar$ exceeds a
    threshold value of $0.25$ (ROIs 2-4) or $0.10$ (ROIs 5, 6) in each simulation setup.
    Note that there is no bar for $D_1$ in ROI 6, because the threshold value was never exceeded.}
    \label{fig:fig2}
\end{figure}

To further quantify the solute transport, we examine the mean
concentrations $\cbar_i(t)$ in six ROIs over time
(\Crefpanels{fig:fig2}{e--g}) and, in addition, the times $\hat{t}_i$
when the mean concentrations first exceed a threshold value
$\hat{c}_i$ (\Crefpanels{fig:fig2}{h}).  Higher diffusion coefficients
result in more rapid transport (\Crefpanels{fig:fig2}{e--g}), with
lower times-to-threshold in ROIs 2--6 (\Crefpanels{fig:fig2}{h}). For
all three diffusion coefficients, the solute first spreads within the
middle ventricle, covering ROIs 2--4, before appearing in the anterior
ventricle (ROI 5) and finally reaching the posterior ventricle
(ROI 6). Notably, the extracellular vesicles, associated with the smallest diffusion
coefficient, do not reach the posterior ventricle (ROI 6) within the
simulated time frame~\Crefpanels{fig:fig2}{b, e}.

\subsection{Model validation by comparison with photoconversion experiments}

To validate the model predictions, we performed photoconversion
experiments in transgenic zebrafish larvae expressing the secreted
Dendra2 in their brain ventricles. The approach is inspired by earlier work
where a photoconvertible protein was injected into the ventricle~\cite{fame2016directional}.
By using a transgenic system where the protein is directly secreted into the CSF,
we avoid detrimental effects of intraventricular injection on CSF properties. 
Our photoconversion protocol
consisted of the acquisition of a baseline fluorescence intensity
$F_0$, before localized exposure to a UV laser (at ROI 1) that
converts Dendra2 from a green- to a red-emitting fluorescent
protein. To quantify the transport of photoconverted proteins, we
measured and calculated the change in fluorescence intensity $\Delta F(t)
= (F(t)-F_0)/F_0$ over time (\Crefpanels{fig:fig3}{a}).

Comparison of the intensity changes with the simulated concentration
profiles in ROIs 1--6 (\Crefpanels{fig:fig3}{a, b}) shows that the
model predictions agree well with the experimental measurements. We
note minor differences however. The simulated values of $\cbar_i$ at
the final time (0.80, 0.76, 0.86, 0.60 and 0.38 in ROIs 2--6, respectively)
are generally higher than the corresponding experimental $\Delta F$
values (0.63, 0.71, 0.67, 0.29 and 0.18). Moreover, we observe
differences in the distribution within the middle ventricle,
especially with regard to ROIs 2 and 3. Overall, we note that the time
to reach a 25\% threshold agree well between the model predictions and
experimental values (\Crefpanels{fig:fig3}{c, d}).

\begin{figure}[H]
    \centering
    \includegraphics[width=0.975\textwidth]{graphics/figure3_compare_experiments_and_simulations.png}
    \caption{
    \textbf{a.} Relative change in fluorescence intensity $\Delta F(t)$ 
    in each region of interest (ROI)
    observed during experiments monitoring
    the movement of Dendra2 proteins photoconverted by laser targeting ROI 1. 
    Bold lines are mean values of the cohort ($n=8$) and shaded regions cover
    one standard devation errors in the mean.
    \textbf{b.} Mean concentration $\cbar(t)$ in each ROI in a
    simulation of Dendra2 protein transport with
    the diffusion coefficient $D_3$.
    \textbf{c.} Time to $\Delta F(t)$ reaches a threshold value of 0.25 in
    one representative zebrafish larva.
    \textbf{d.} Time to $\cbar(t)$ reaches a threshold value of 0.25 for
    a transport simulation with diffusion coefficient $D_3$.
    We have excluded the posterior-most regions of the geometries,
    because the values of $\Delta F$ and $\cbar$ never exceeded the threshold
    value of 0.25 in these regions. Scale bars 50 \textmu m.}
    \label{fig:fig3}
\end{figure}
% Final Delta F values
%0.6322650156762044 ROI 2
%0.7059449124175479 ROI 3
%0.6694500367160006 ROI 4
%0.28986700094045437 ROI 5
%0.18155055624229066 ROI 6
% Final mean c values ROIs 2-6: 0.80, 0.76, 0.86, 0.60 and 0.38

\subsection{Absence of ciliary motion affects local solute distribution}

Ciliary motion is integral to physiological CSF flow, and absence of
cilia has been associated with defective brain development and various
pathologies~\cite{Eichele2020Cilia-drivenVentricle,
  Olstad2019CiliaryDevelopment, Faubel2016Cilia-basedVentricles,
  Guirao2010CouplingCilia, Afzelius2004CiliaRelatedDiseases,
  Sawamoto2006NewBrain, Yoshiba2014RolesSymmetry,
  Hirokawa2006NodalAsymmetry}.  As observed both experimentally and
computationally, absence of motile cilia alters CSF flow patterns.  We
next investigate the implications of these changes on intraventricular
transport. We experimentally measured the movement of proteins by
monitoring photoconverted Dendra2 signal in control zebrafish and
motile-cilia mutant \emph{schmalhans} (\emph{smh}) zebrafish.  In ROIs
2 and 4, which quantify Dendra2 distribution downstream of the
photoconversion site, we observed a slower increase of fluorescence
intensity (\Crefpanels{fig:fig4}{a}): the mean times-to-threshold in
the \emph{smh} mutants were $45\%$ (ROI 2) and $25\%$ (ROI 4) higher
compared to controls (\Crefpanels{fig:fig4}{b}). We found no
significant differences between \emph{smh} mutants and controls in the
Dendra2 distribution in ROIs 3, 5, and 6.

In the computational model, we represent the \emph{smh} mutants by
considering transport of Dendra2 with CSF flow driven by cardiac
pressure pulsations only, representing the absence of motile
cilia. Comparing the in silico mutants with transport predictions in
the presence of both cardiac pulsations and motile cilia, we observe
noticeably delayed dynamics in ROIs 2, 4 and 5, but little change for
ROIs 3 and 6 (\Crefpanels{fig:fig4}{c}). Compared to experimental
data, the control and mutant scenarios differed more for the
simulations, with the times-to-threshold in ROI 2 and 4 being 81\% and
28\% higher in the absence of cilia-mediated flow. Moreover, the
predicted mean concentrations in ROI 5 deviate between the two
scenarios, which is not the case in the experiments. Altogether, our
findings identify that, for smaller proteins such as Dendra2,
diffusion plays an important role for solute movement to regions
upstream in the cilia-mediated flow fields or to distant ventricles,
while the cilia affect the downstream mean concentrations.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure4_compare_control_mutant.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $\Delta F(t)$
    in each region of interest (ROI) for control (black lines, $n=8$)
    and \emph{smh} mutant (red lines, $n=6$) fish. 
    Bold lines are mean values of the cohorts and shaded regions cover one standard devation errors
    in the means.
    \textbf{b.} Times when $\Delta F(t)$ first exceeds a threshold value,
    equal to 0.25 (ROIs 2--4) or 0.10 (ROIs 5, 6), for control (gray dots) and
    \emph{smh} mutant (red dots) fish.
    The thick black lines denote mean values of the datasets. The $p$-values
    were calculated using a Mann-Whitney U test with a confidence level of 95\%, with 
    the null hypothesis that the two distributions of control and mutant data are 
    from the same population.
    \textbf{c.} Mean concentration $\cbar(t)$ in each ROI,
    simulated using the baseline model (black lines) and the cardiac-induced/no-cilia model (red lines).
    The horizontal gray lines mark the threshold value $\hat{c}$, and the time
    $\hat{t}$ when the threshold is first exceeded is reported in the plot legends.
    Note that in both panels \textbf{a} and \textbf{c}, the vertical axes have
    different limits in each plot.}
    \label{fig:fig4}
\end{figure}

\subsection{Regional loss of cilia motility delays transport}

We previously characterized three distinct ciliated cell lineages in
the zebrafish larval brain~\cite{Olstad2019CiliaryDevelopment,
  DGama2025MotileBrain}.  To identify the effects of these lineages on
CSF flow and transport, we investigate region-specific cilia paralysis
with the computational model.  Three cilia-covered regions were
paralyzed separately: the anterior ventricle, and the dorsal and
ventral parts of the middle ventricle (\Crefpanels{fig:fig5}{a}).  To
simulate regional cilia paralysis, we remove the tangential traction exerted by
the cilia in the respective region.
%
%% the cilia tangential traction boundary condition is for each case
%% changed to a free-slip boundary condition in the paralyzed region,
%% prescribing zero tangential forces to represent absence of cilia
%% motion.
%

Removing cilia in the dorsal part of the middle ventricle has a strong
effect on the CSF velocities: the maximum velocity magnitude drops
from 27.9 \textmu m/s to 7.9 \textmu m/s. In vivo, these dorsal cilia
generate the strongest flow at this developmental stage of the
fish~(\Crefpanels{fig:fig1}{c})~\cite{Olstad2019CiliaryDevelopment}.
In contrast, only removing cilia in the anterior ventricle or the
ventral region of the middle ventricle has minimal impact, and results
in maximum speeds of 27.9 \textmu m/s and 27.4 \textmu m/s,
respectively.  We remark that out of the total tangential forces,
ventral cilia in the middle ventricle contribute by 34\%, while the
dorsal cilia exert 52\% of the total tangential
force~(\Crefpanels{fig:fig5}{b}). \mer{Thus, the CSF velocities are
  affected not only by the magnitude of the applied forces, but result
  from a nontrivial interplay between force, directionality and
  location.}
%% The magnitude of these force
%% contributions are thus more similar than the difference in their
%% impact on the maximum velocity magnitudes.

Turning to transport of Dendra2, we again note that removal of the
dorsal cilia has the most pronounced effect. All mean concentration
profiles (except ROI 1, where we impose the photoconversion curve)
change noticeably (\Crefpanels{fig:fig5}{c}). Compared to the baseline
model results, we observe less solute distribution in the middle and
anterior ventricles (ROIs 2--5), and more rapid transport to the
posterior ventricle (ROI 6). On the other hand, removal of anterior
cilia does not alter the mean concentrations
(\Crefpanels{fig:fig5}{c}). Finally, without ventral cilia in the
middle ventricle, more solute spreads towards the anterior ventricle
(ROI 5) and less solute spreads towards the posterior ventricle (ROI
6).
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure5_compare_cilia_modifications.png}
    \caption{\textbf{a.} The three separate cilia regions for which cilia paralysis is simulated
    by imposing a zero tangential traction boundary condition in the respective region.
    \textbf{b.} The total tangential force applied to the boundary with the tangential traction
    boundary condition in each of the three separate ciliated regions, as well as the total force
    (the sum of the three regions), calculated with \Cref{eq:tangential_force_calculation}. 
    \textbf{c.} Mean concentration $\cbar(t)$ for the six regions of interest (ROIs) in the three 
    paralysis scenarios considered. The region named in the legend is the region
    where we modify the cilia tangential traction boundary condition to a free-slip condition
    in the CSF flow model.}
    \label{fig:fig5}
\end{figure}

\subsection{Modification of ventricular geometry impacts solute distribution}
Variations in ventricular morphology impact CSF flow
patterns~\cite{Olstad2019CiliaryDevelopment}.  To assess how these
variations influence solute distribution, we consider four alternative
ventricular geometries: (i) constriction of the anterior-middle
ventricular duct (66 \% reduction in cross-sectional area);
(ii) constriction of the middle-posterior duct (33 \% reduction);
(iii) middle ventricle shrunk in the lateral direction (43 \% reduction);
and (iv) all three combined (\Crefpanels{fig:fig6}{a, b}).
%
%% one where the duct connecting the anterior and middle ventricles is constricted
%% (66 \% reduction in cross-sectional area);
%% one where the duct connecting the middle and posterior ventricles is constricted
%% (33 \% reduction);
%% one where the middle ventricle is shrunk in the lateral direction (43 \% reduction);
%% and the combination of these three modifications (\Crefpanels{fig:fig6}{a, b}).
%
%
In these four modified geometries, the total volume is reduced by
3.9\%, 4.1\%, 1.4\% and 9.4\%, respectively, compared to the original
geometry. \mer{MER: HH adds a sentence on changes in tangential
  traction(\Crefpanels{fig:fig6}{d}). This is notwithstanding a
9.5\% reduction in ventral cilia force magnitude
(\Crefpanels{fig:fig6}{d}), which are the cilia that propel CSF and solute towards ROI 6.)}

Constricting the connection between the anterior and middle ventricle
reduces solute transport to the anterior ventricle (ROI 5), and the
mean concentration in ROIs 2--4 and 6 slightly increases, compared to
the original geometry (\Crefpanels{fig:fig6}{c}). \mer{We thus note
  that restricting the ...} Constricting the middle-posterior duct
mainly delays the transport into the posterior ventricle (ROI 6),
while \mer{gently} accelerating transport towards the anterior
ventricle (ROI 5). Faster transport to the anterior ventricle (ROI 5)
and slower transport to the lower middle and posterior ventricle
(ROI3, ROI 6) is also the case for the geometry with a shrunk middle
ventricle. \mer{Interestingly, this reduction in the transport to the
  posterior ventricle (ROI 6) happens in spite of a 6.6\% increase in
  the ventral cilia forces.}
%
%For the mesh where all three ventricles are modified,
When combining all three modifications,
we observe higher retention of solute in the middle ventricle (ROIs 2--4),
while the mean concentration is reduced in both the anterior
and the posterior ventricles (ROIs 5 and 6)s.

To summarize, these results show that the ventricular geometry and the
cross-sectional area of the ventricular ducts clearly impact solute
transport, with notable differences in transport resulting from modest
geometry modifications. These findings may explain the larger
differences in the cilia-deficient computational model results
compared to the experimental data in \Cref{fig:fig4}, \mer{as the
computational geometry is not an exact representation for the
individual zebrafish ventricles.}

\lyng{We previously identified large interindividual variation in ventricular size and
CSF velocities between larva at given developmental stages and throughout
development~\cite{Olstad2019CiliaryDevelopment, DGama2025MotileBrain, Jeong2024TheZebrafish}.
Our results suggest that these two parameters, CSF velocities and ventricular size,
affect each other.}
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{graphics/figure6_compare_modified_geometries.png}
    \caption{\textbf{a.} The original computational mesh of the ventricular geometry.
    \textbf{b.} The four modified versions of the computational geometry. Compared
    to the original mesh, the volumes of the deformed meshes are reduced by 3.9\%, 4.1\%, 1.4\% 
    and 9.4\%.
    \textbf{c.} Mean concentration $\cbar(t)$ simulated with the original and modified geometries
    in the 6 ROIs.
    \textbf{d.} A comparison of the tangential cilia force in the three cilia regions
    and the total force applied, calculated with \Cref{eq:tangential_force_calculation}.
    In each region, the force $F_{\mathrm{original}}$ 
    applied on the original mesh is the baseline of 100\%,
    to which the forces $F_i$ of the other mesh versions are compared, where $i$ denotes
    the cases fore-shrunk, middle shrunk, mid-hind shrunk, and the combination of the three.
    The bars for the four shrunk scenarios display $F_i/F_{\mathrm{original}}\times 100\%$.}
    \label{fig:fig6}
\end{figure}

\section{Discussion}
In this study, we have presented a computational model of CSF flow and
solute transport within brain ventricles.
We used the larval zebrafish brain ventricles since its geometry and
flow mechanisms are well described, thereby allowing us to validate
our simulations experimentally. Our computational framework can be adapted to
and implemented in other ventricular systems in the future,
once the flow parameters of such systems
are adequately resolved.
We assumed that intraventricular CSF flow was governed
by the Stokes equations.
By imposing tangential traction and normal pressure
on the CSF flow, we were able to reproduce flow fields with similar features
as observed experimentally. The prominent flow features were vortex-structured flow
mediated by cilia, and pulsatile flow mediated by the cardiac cycle.

Using the simulated CSF flow, we modeled solute transport by an advection-diffusion equation.
The transport results compared well with data from in vivo experiments in zebrafish.
Then, by using a computational model as well as zebrafish mutants with paralyzed cilia,
we identified that even though advection by cilia is dominant,
diffusion plays a significant role in protein transport.
Finally, we investigated how local cilia paralysis and ventricular geometries
affect solute distribution. Altogether, we present not only new findings on the role of
diffusion and advection in intraventricular transport, but also provide a
computational framework to study CSF flow and transport in ventricular geometries.

\subsection{The roles of advection and diffusion in intraventricular solute transport}
To investigate the roles of advection and diffusion in solute transport, we quantified transport
based on changes in fluorescence intensity $\Delta F$ (experiments) and
mean concentration $\cbar$ (simulations) over time. Our results convincingly show that
intraventricular transport of solute depends on a balance of advection and diffusion.
While advection is the strongest transport mechanism on a global scale for
Dendra2 proteins (global Péclet number 13), diffusion is balancing out
advection for shorter distances and for regions without motile cilia activity.
For instance, for the larger extracellular vesicles, no solute reached the (distant)
posterior ventricle (a distance of $L = 200$ \textmu m) within the simulation time (around 860 s).
Advective transport by the mean velocity $\overline{U}=2.4$ \textmu m/s over the same
distance would happen on a timescale of $t_a=L/\overline{U}=83$ s, indicating that
transport to the posterior ventricle is solely driven by diffusion at the developmental
stage 2 dpf. The estimated timescale of diffusion $t_d=L^2/D_1$ over the 
same distance is $\approx$24500 s. Larger molecules
transported primarily by advection are thus more easily
compartmentalized within the ventricles
as a result of the compartmentalized CSF flow.

Removing motile cilia slowed down transport within
the middle ventricle and towards the anterior ventricle,
but the transport towards the posterior ventricle was unaffected.
Photoconversion experiments on cilia mutant
zebrafish larva (\emph{smh}) aligned with these observations.
This further indicates that
transport from the middle to the posterior ventricle is
purely driven by diffusion at this developmental stage.
The experimental results provide not only confidence
in the predictive ability of the computational model,
but also quantify the impact of cilia motility
on intraventricular transport.
Interestingly, Yoshida~\emph{et al.} observed a drastic reduction
in CSF mixing in specific regions of the brain
ventricles when simulating reduced cilia motility~\cite{Yoshida2022EffectVentricles}.
\lyng{Since impaired brain 
clearance of tracers has been demonstrated in idiopathic normal pressure hydrocephalus
patients~\cite{Eide2020MagneticHydrocephalus},
cilia-enhanced mixing in the brain ventricles might be
fundamental to physiological transport within the brain.}

\subsection{Model validation and choice of parameters}
Even though we identified similar dynamics in simulations and
in vivo experiments, the dynamics of the simulations were generally
more rapid and resulted in higher total amounts of solute.
Differences between experiments and simulations may relate to
ventricular size, as we showed that modifying the geometry of the
ventricular geometry impacts solute distribution,
and the ventricular size varies substantially 
in the control population at 2 dpf.
Additionally, differences may result
from the definition of the photoconversion site (ROI 1)
as the volume of this region determines the total amount of solute
and thereby the rate of transport. Another source
of discrepancy may relate to the permeability of the ventricular wall,
which we have considered as impermeable in our simulations.
The neuroepithelium lining the walls of brain ventricles in 24 hour old zebrafish
is permeable to dyes smaller than 70 kDa~\cite{Chang2012AnNeuroepithelium}.
Since Dendra2 is around 26 kDa in size~\cite{Gurskaya2006EngineeringLight} and
thus under this reported limit, one could assume that Dendra2 may
leak out of the brain ventricles over the relevant time spans
and render the dynamics slower. 
It may also be the case that our model predicts faster
transport because the Dendra2 diffusion
coefficient we used (based on an experimental value
\cite{GuraSadovsky2017MeasurementExpansion}) is too high.
This could be because embryonic CSF has higher viscosity
than the solution considered in the measurements, 
or the presence of proteins, nutrients and other solutes
that hinder diffusive transport
\cite{Bloomfield1998EffectsFluid, Einstein1905UberTeilchen, Cussler2009Diffusion:Systems}.
Finally, an important difference
is the three-dimensional nature of the computational model,
while the experimental images only show 2D slices.

We modeled motile cilia as a tangential
traction at the ventricular walls
while allowing the CSF to slip,
motivated by experimental observations of
flow fields in zebrafish larval brain ventricles in which
the fluid velocity increases towards the
wall~(\Crefpanels{fig:fig1}c)~\cite{Olstad2019CiliaryDevelopment}. 
The ciliated boundaries in our model are regionally homogeneous,
whereas in reality, cilia populate the ventricle walls in a
heterogeneous manner~\cite{Olstad2019CiliaryDevelopment}. Heterogeneity
in cilia location, beating direction and beating frequency
has been shown to enhance particle clearance in airway cilia arrays
in mice~\cite{Ramirez-SanJuan2020Multi-scaleArrays}.
Other studies have demonstrated that changes to beating directionality affects transport,
and that specific coordination of the ciliary beating is important in the brain ventricles for
flow and signaling processes~\cite{Eichele2020Cilia-drivenVentricle, Olstad2019CiliaryDevelopment,
Faubel2016Cilia-basedVentricles, Guirao2010CouplingCilia, Afzelius2004CiliaRelatedDiseases,
Pellicciotta2020Cilia},
for maintaining healthy functioning of mucous transport in
airways~\cite{Rayner1996CiliarySyndrome, Schneiter2021Multi-scaleFunction,
Bustamante-Marin2019LackClearance, Tsukita2012CoordinatedFeet},
and in the nodal flow that establishes
left-right symmetry in mammals~\cite{Yoshiba2014RolesSymmetry, Hirokawa2006NodalAsymmetry,
Sawamoto2006NewBrain}.
Incorporating spatial heterogeneity in our model would
therefore presumably affect results, and could be
interesting to pursue in further work.

The maximum stress
$\vert\btau\vert_{\mathrm{max}} = 2.5 \ \mathrm{mPa} = \num{2.5e-3} \ \mathrm{N/m^2}$
of the cilia boundary condition
was applied in the dorsal-posterior region of the middle ventricle.
In a model of CSF flow in human brain ventricles,
Siyahhan~\emph{et al.} modeled ciliary motion with
a force that increased linearly from
the ventricular wall to a maximum force density of
$f_{\mathrm{max}} = 526$ N/$\mathrm{m^3}$ at a distance 15 \textmu m (the cilium length)
away~\cite{Siyahhan2014FlowVentricles}.
In a cilia array of thickness $\delta = 15$ \textmu m,
the maximum stress $\vert\btau\vert_{\mathrm{max}}$
in our model would correspond to a maximum force density of 
\begin{equation*}
    f_c = \frac{\vert\btau\vert_{\mathrm{max}}}{\delta}
        = \frac{\num{2.5e-3} \ \mathrm{N/m^2}}{\num{15e-6} \ \mathrm{m}}
        = 169 \ \mathrm{N/m^3}.
\end{equation*}
With $\delta=5$ \textmu m (the typical cilium length in zebrafish brain ventricles
\cite{Olstad2019CiliaryDevelopment,
Salman2022ComputationalEmbryo, Ringers2023NovelEpithelia}) 
our cilia stress corresponds to a maximum force density of
$f_c=507$ N/$\mathrm{m^3}$.
For further comparison, Thouvenin~\emph{et al.}
modeled two-dimensional cilia-mediated 
flow in the central canal of zebrafish
embryos using a force density of 4000 N/$\mathrm{m^2}$\cite{Thouvenin2020OriginCanal}.
Averaging this value over the reported central canal diameter 8.9 \textmu m
yields a maximum force of 449 N/$\mathrm{m^3}$.
It is very interesting to note the similarity 
in cilia force densities estimated in our and these two previous studies.
\fixme{Previous work concluded that motile cilia align by response to the shear
stress of flow. Perhaps there is an optimal cilia stress?}

To simulate protein transport in the brain ventricles we solved
an advection-diffusion equation. We handled advective transport
through the anterior/posterior boundaries
(where cardiac pressure pulsations are imposed)
with an approximate periodic boundary condition
that conserves the total solute concentration.
A possible improvement to this would be
to consider a 3D-0D coupling of in- and outflow using
ordinary differential equations,
similar to applications in cardiac
modeling~\cite{Brown2024AMechanics, Augustin2021ACirculation}.
Alternatively, the beating cardiac motion could be modeled by ventricular wall deformation
to avoid advective boundary fluxes~\cite{Causemann2022HumanFramework,
Kurtcuoglu2005ComputationalSystem, Kurtcuoglu2007ComputationalSylvius, Linninger2005PulsatileBrain}.

% The Stokes-Einstein relation assumes the diffusing
% substance is of spherical shape and the relation does not
% necessarily hold in CSF. Moreover, the extrapolation implies the assumption that
% the two substances (the one for which we determine $D$ and the one we extrapolate from)
% have equal densities. We note that one previous study reported that
% Dendra2 satisfies the Stokes-Einstein in a solution of similar viscosity
% to CSF\cite{GuraSadovsky2017MeasurementExpansion}.
% We used CSF density and viscosity values measured in
% humans~\cite{Bloomfield1998EffectsFluid},
% and because embryonic CSF in zebrafish is high in proteins, nutrients and other substances,
% the CSF in which the photoconverted Dendra2 are transported in our experiments may be more viscous
% than we assumed. This could reduce the diffusion coefficient (cf.~\Cref{eq:D_stokes_einstein}),
% but previous works have reported a limited impact of protein concentration on the viscosity of
% CSF~\cite{Bloomfield1998EffectsFluid, Brydon1995PhysicalViscosity}.
% Merely the presence of other proteins and nutrients may also reduce the diffusion
% coefficient by reduction of mean-free path lengths of molecules~\cite{Cussler2009Diffusion:Systems}.

\subsection{Numerical approximation}
The Brezzi-Douglas-Marini and discontinuous Galerkin scheme
$\mathrm{BDM}_1$-$\mathrm{DG}_0$
we used to discretize the Stokes equations
is computationally more expensive
than \fixme{conventional} methods such as a
Taylor-Hood discretization
($P_k-P_{k-1}$ elements)~\cite{Stenberg1990ErrorProblem} or lower-order enriched
elements (such as $P_1-P_1$ with a velocity stabilization)~\cite{Brezzi2011MixedMethods}.
Here, $P_k$ is the space of continuous Lagrange polynomials of degree less
than or equal to the integer $k$.
We opted for the BDM-DG scheme because of its exactly-mass-conserving property;
by the definition of the elements, the numerical approximation of
the CSF velocity $\uu_h$ satisfies the continuity equation (\Cref{eq:stokes_eqs_cont}).
This yields $\normlinf{\nabla\cdot\uu_h}=0$,
an important property for the stability of the advection-diffusion
equation that we used to simulate solute transport,
especially in advection-dominated cases.
\cite{cesmelioglu2022compatible, johnson2009numerical, gresho1998incompressible}.
Additionally, the BDM-DG scheme 
lets us impose the impermeability condition $\uu\cdot\nn=0$
strongly, since the BDM elements control the normal component of the 
velocity~\cite{Brezzi1985TwoProblems}.
For elements where this is not the case (such as Taylor-Hood elements),
the impermeability boundary condition must
be handled with for example a Nitsche method~\cite{Nitsche1971UberSind}
or with a Lagrange multiplier~\cite{Babuska1973TheMultipliers,
Bertoluzza2017BoundaryHemodynamics}, which affects
convergence properties and increases problem complexity
(see \Cref{subsec:appendixA1} for more details).

\subsection{Outlook}
We studied the zebrafish larval brain ventricles,
in which the flow generated by ciliary motion
is of the scale of the geometry~\cite{Olstad2019CiliaryDevelopment}.
In large brains such as the human brain, ciliary beating mostly 
contributes to flow in localized regions proximal to the ventricular walls
due to the microscopic length of the cilia~\cite{Siyahhan2014FlowVentricles}.
The cilia may play an important role in regulating the
amount of substances like extracellular vesicles close to the ventricular walls,
similar to coral reefs,
where cilia are primary regulators of
oxygen amounts at the coral surface~\cite{Pacherres2022CiliaryProduction}.
Flow mediated by cilia indeed serves such a purpose in nodal flow at early
developmental stages of the brain~\cite{Hirokawa2006NodalAsymmetry}, and the discovery
of a cilia-driven flow network in rodent brains suggests such selective
brain ventricular transport~\cite{Faubel2016Cilia-basedVentricles}.
Our simulation studies demonstrate that our computational model is
flexible and allows in silico testing of such open questions
and further hypotheses on the effects of ciliated
lineages and ventricular morphology on CSF flow and intraventricular transport.

% Small versus large brain: For instance, a large
% volume would require an extensive amount of time for solutes to distribute
% if based solely on diffusion. Besides, the presence of multiple cilia
% instead of a single cilium per cell -- the former can be found in
% mammalian ependymal cells -- will generate stronger velocities.
% Hence advection may be more dominant in those situations.

% Much evidence indicate that blood vessel pulsations, respiration, CSF secretion
% and brain motion induce the pulsatile bulk flow in the central
% regions of the ventricles~\cite{Vinje2019RespiratoryMeasurements,
% Siyahhan2014FlowVentricles, Linninger2005PulsatileBrain, Enzmann1992BrainImaging,
% Kurtcuoglu2007ComputationalSylvius}.

% To date, motile-cilia deficiency has been associated with altered brain function
% in zebrafish, dysfunctional establishment of left-right symmetry,
% and ventricular enlargement in zebrafish, rodents and humans.~\cite{Afzelius2004CiliaRelatedDiseases,
% Youn2018PrimaryDiseases, Ma2024CiliaDisease, Ibanez-Tallon2002LossHydrocephalus}

\section{Methods}

\subsection{Zebrafish maintenance, genotyping and strains in photoconversion experiments}
The animal facilities for zebrafish~(\emph{danio rerio}) are approved by the
Norwegian Food Safety Authority (NFSA, Mattilsynet).
The zebrafish were maintained in accordance with the guidelines set by the
NFSA and the European Communities Council Directive.

The larval and adult zebrafish were raised under standard husbandry
conditions at 28.5°C in a Techniplast Zebtech Multilinking system.
The fish tanks were kept at constant pH 7.0 and 685 \textmu S with a
14/10 hr light/dark cycle.
From fertilization to 3 days post-fertilization (dpf), larvae were maintained in egg
water (1.2 g marine salt and 0.1\% methylene blue in 20 L reverse osmosis (RO) water)
and subsequently transferred to artificial fish water (AFW)
(1.2 g marine salt in 20 L RO water). The zebrafish lines used in the experiments were
\emph{ccdc103(dnaaf19)\textsuperscript{tn222a}}
(\emph{schmalhans (smh)})~\cite{Jau-NianChen1997Left-rightZebrafish}
and \emph{Tg(gfap:Gal4FF)\textsuperscript{nw7Tg}}~\cite{DiazVerdugo2019Glia-neuronSeizures},
\emph{Tg(5xUAS:Signal-Dendra2)\textsuperscript{nw21Tg}}.
Animals were in the pigmentless \emph{nacre\textsuperscript{b692}}
(\emph{mitfa\textsuperscript{-/-}})~\cite{JamesA.Lister1999NacreFate} background.

Wholemount in vivo live imaging and photoconversion
experiments were performed with zebrafish larvae at the 2 dpf stage obtained from inbreeding of heterozygous
\emph{smh\textsuperscript{+/-};Tg(gfap:Gal4FF);Tg(5xUAS:Signal-Dendra2})
adult animals, as described by D'Gama~\emph{et al.}~\cite{DGama2024Cilia-mediatedBrain}.
Controls consisted of either wild-type (\emph{smh}\textsuperscript{+/+})
or \emph{smh}\textsuperscript{+/-} heterozygous zebrafish from the same breeding.
The mutants were identified based on their curved body~\cite{Jau-NianChen1997Left-rightZebrafish}.

For genotyping, genomic DNA (gDNA) was isolated from clipped fins of anesthetized adult
fish using 100 \textmu L of PCR lysis buffer (containing 1M tris pH 7--9, 0.5 M EDTA,
tritonX-100, and Proteinase K 0.1 mg/ml) overnight at 50°C.
To stop the lysis reaction, the samples were heated to 95°C for 10 minutes
and then centrifuged at 13000 rpm for 2 minutes.
The supernatant containing gDNA was utilized for KASP assays-based analysis.
The gDNAs were diluted (1:2) with water. 
The KASP assay was run following the guidelines of the
manufacturer (LGC Biosearch Technologies™) with 3 \textmu L gDNA,
5 \textmu L of master mix, 0.14 \textmu L of assay mix
and 1.86 \textmu L of RO water per well on a 96-well plate.

\subsection{Generation of transgenic line to study CSF and transport dynamics} 
To generate a transgenic line expressing the photoconvertible protein Dendra2 in
the brain ventricles, we fused the open reading frame (ORF)
of the neuropeptide y (npy) signal peptide
(\emph{npy}: ENSDARG00000036222, Q1LW93) to the N-terminal of
zebrafish codon optimized Dendra2 DNA sequence.
The DNA sequence was synthesized with EcoR1 enzyme sites at the 5’ and
3’ ends (GenScript Biotech) and inserted into the
pT2MUASMCS vector~\cite{Asakawa2008GeneticZebrafish}
through restriction enzyme cloning (GenScript Biotech)
to generate the 5xuas:Signal-Dendra2 plasmid DNA.

A volume of 2 nl of a mixture of the plasmid DNA (60 pg) and tol2 mRNA (10 pg)
was microinjected into one-cell stage embryos to generate the transgenic line,
as described in Jeong~\emph{et al.}~\cite{Jeong2024TheZebrafish}. The injected
embryos were raised to adulthood (F0).
Germline-transmitted founder zebrafish were identified by breeding with
multiple Gal4 transgenic lines. Stable F1 embryos expressing the Gal4-driven
Dendra2 signals were screened and raised to adult zebrafish.

\subsection{Wholemount zebrafish in vivo time-lapse imaging and photoconversion of Dendra2}
Two dpf larval zebrafish were anesthetized in 0.013\% MS222
(Ethyl 3-aminobenzoate methanesulfonate, Sigma) in AFW and
mounted laterally in 1.5\% low melting agarose in a Flurodish (VWR, FD35PDL-100),
to obtain a lateral view of the brain ventricles. After the
fish were positioned in the agarose, the dish stood for 5 min to allow the agarose to solidify.
After solidification, AFW containing 0.013\% MS222 was added,
and then the fish were transferred to the confocal microscope (LSM880 Examiner, Zeiss)
and imaged using a 20x water-immersion objective (Zeiss, NA 1.0, Plan-Apochromat)
at room temperature.

The time-lapse images were acquired in a single plane, covering the
telencephalic, diencephalic and rhombencephalic ventricles simultaneously,
with a frequency of 0.37--0.38 Hz (2.60--2.68 sec / frame). The size
of the images was $1024\times400$ pixels. In total, 300 images were acquired per fish.
The first 10 images were scanned without photoconversion to obtain a
baseline value of fluorescence intensity, and the next 290 images were
scanned while performing photoconversion.

For photoconversion, ``Bleaching" and ``Region" in the ZEN software
were used. A 405 nm laser was focused into a circular area (diameter: 16 \textmu m, scan 0.77 \textmu sec/pixel)
of the dorsoposterior diencephalic ventricle with 100\% laser power.
The laser was illuminated in the area repeatedly between each scanning.
After imaging, the fish health was checked and only data from healthy fish were analyzed.
Data were collected and analyzed from three separate experiments,
with a total of 8 controls and 6 mutants.

\subsection{Photoconversion data analysis}
The image data analysis was performed in MATLAB and the codes are
openly available~\fixme{\cite{zenodo-link-to-all-code}}.  The
photoconverted channel was aligned to correct for drift in $x,y$
directions using previously developed
codes~\cite{Reiten2017Motile-Cilia-MediatedComputations,
  Ringers2023NovelEpithelia}.  Only stable recordings were further
analyzed.  To increase the signal-to-noise ratio, the images were
downsampled by a block average using a resampling factor of 6. 
The change in fluorescence intensity
$\Delta F = (F-F_0)/F_0$ was
calculated for each pixel and each time point.  Here, the
value $F_0$ is the average fluorescence intensity for the baseline
(the first 10 time-lapse images) acquired before photoconversion.

To identify the time needed to pass a certain threshold value,
$\Delta F$ values were first smoothened.
The first timepoint when values surpassed the threshold was reported
per pixel (in the downsampled images).
A mask for the image was generated using the green (not photoconverted)
channel and calculated on the block-averaged data based on the intensity
being higher than 1.5x median intensity of all the frames.

To obtain $\Delta F$ values for each region of interest (ROI), ROIs were
first drawn on the aligned and block-averaged time series (\Cref{fig:fig2}).
Six ROIs were drawn manually: ROI 1 around the location of photoconversion,
and ROIs 2--6 in different regions of the ventricular system.
The pixel values of the fluorescence intensity within one ROI were
averaged for each time point, and the relative change $\Delta F$ calculated with this averaged value.
To normalize the data with respect to photoconversion efficiency, the
$\Delta F$ curves for all ROIs were divided by the $\Delta F$ values
obtained for ROI 1 at the end of the experiment,
so that $\Delta F$ values for the photoconverted site ROI 1 ranged from
zero to one. Finally, we averaged the $\Delta F$ data from controls and from mutants.

\subsection{Intraventricular injection of microbeads and particle tracking}
We carried out particle tracking in a larval zebrafish injected with fluorescent beads,
as described in previous work~\cite{Olstad2019CiliaryDevelopment}.
Anesthetized 2 dpf zebrafish were injected with 1 nl of a mixture containing 0.1\% w/v fluorescent beads
(SPHERO Fluorescent Yellow Particles 1\% w/v, F = 0.16 mm) diluted
in 7.5 mg/ml 70 kDa rhodamine B isothiocyanate-dextran (RITCdextran; Sigma-Aldrich, R9379)
dissolved in artificial CSF. Artificial CSF composition was as follows:
124 mM NaCl, 22 mM D-(+)-Glucose, 2.0 mM KCl, 1.6 mM MgSO\textsubscript{4} $\cdot$ 7 H\textsubscript{2}O,
1.3 mM KH\textsubscript{2}PO\textsubscript{4}, 24 mM NaHCO\textsubscript{3},
2.0 mM CaCl\textsubscript{2} $\cdot$ 2 H\textsubscript{2}O.
The needles used for the injections were pulled with a
Sutter Instrument Co. Model P-2000 from thin-walled glass
capillaries (1.00 mm; VWR) and cut open with a forceps.
A volume of 1 nl of the solution was injected with a pressure injector
(Eppendorf Femtojet 4i) in the rostral rhombencephalic ventricle.
The pressure and time were calibrated for each needle using a 0.01 mm calibration slide.

Following injection, the zebrafish was directly placed under the confocal
microscope and 1200 images were acquired at a frequency of 13.1 Hz.
A single optical section was obtained. Particle tracking was performed using
TrackMate~\cite{Tinevez2017TrackMate:Tracking} in Fiji/ImageJ~\cite{Schindelin2012Fiji:Analysis}
and plotted using MATLAB, as described by Jeong~\emph{et al.}~\cite{Jeong2022MeasurementTelencephalon}.
The parameters for TrackMate were as follows.
The DoG detector was used with the settings: threshold: 2.0, with
median filtering; radius: 2.0, with subpixel localization.
Next, the Simple LAP tracker was used with settings: max frame gap: 2;
alternative linking cost factor: 1.05; linking max distance: 5.0; gap closing max distance: 2.0;
splitting max distance: 15.0; allow gap closing: true; allow track splitting:
false; allow track merging: false; merging max distance: 15.0; cutoff percentile: 0.9.
Only the particle tracks with at least 30 data points, while simultaneously
covering a distance of at least 8.8 \textmu m, are shown (\Crefpanels{fig:fig1}{b, c}).

\subsection{Image-based computational geometries of zebrafish brain ventricles}
To model CSF flow and transport in the zebrafish brain ventricles, we
generated a 3D surface representation of the ventricular walls using
confocal imaging data of a zebrafish embryo injected
intraventricularly with a 70 kDa dye at 2
dpf~\cite{Olstad2019CiliaryDevelopment}. At 2 dpf, the ventricular
system consists mainly of 3 cavities: the telencephalic (anterior),
the di-/mesencephalic (middle) and the rhombencephalic (posterior)
ventricles, which are connected by interventricular ducts. From the
surface geometry, we constructed a volumetric mesh of the ventricles
by initially using fTetWild~\cite{Hu2020FastWild} and subsequently
local refinement functionality in DOLFINx~\cite{dolfinx2023preprint}. 
The geometry spans approximately 600 microns along the rostrocaudal axis
and around 300 microns in the lateral direction. At most around
100 microns separates the ventral and dorsal ventricular walls.
The standard mesh consists of 141 512 tetrahedral cells with a
maximal (minimal) edge length of 14.9 \textmu
m (3.1 \textmu m) (\Crefpanels{fig:fig1}{d}). Parts of the
ventricular surfaces were marked to be lined with motile cilia
(\Crefpanels{fig:fig1}{d}, magenta
markers)~\cite{Olstad2019CiliaryDevelopment}.

The morphology and geometry of the brain ventricles vary across zebrafish
individuals, and under physiological and pathological conditions~\cite{Olstad2019CiliaryDevelopment}. 
Motivated by an interest in how ventricular geometry impacts solute distribution,
we also consider four variations in the geometry (cross-sectional area
reduction in parentheses):
by shrinking the fore-mid brain connection (66 \%), 
the middle ventricle (33 \%), the mid-hind brain connection (43 \%), and all three simultaneously.
We used Blender~\cite{Community2018BlenderPackage} to modify the geometries (\Crefpanels{fig:fig6}{a, b}).
% Cross-sectional areas of ducts normal to flow direction
% Modified region       Original mesh        Deformed mesh
% Fore                  3.36e-3 mm^2         1.15e-3 mm^2
% Mid                   6.11e-3 mm^2         4.08e-3 mm^2
% Hind                  2.91e-3 mm^2         1.67e-3 mm^2

\subsection{Computational CSF dynamics in the brain ventricles}
The beating motion of the cilia generates CSF flow with persistent rotational
structures~\cite{Olstad2019CiliaryDevelopment}
within the brain ventricles (\Crefpanels{fig:fig1}{a--c}).
In addition, cardiac pulsations induce pulsatile CSF flow.
We model this flow of CSF by the time-dependent,
incompressible Stokes equations which read as follows.
Find the CSF velocity $\uu = \uu(\xx, t) = (u_x(\xx, t), u_y(\xx, t), u_z(\xx, t))$
and the CSF pressure $p = p(\xx, t)$ for $\xx \in \Omega$ and time $t>0$, such that
\begin{subequations}
    \begin{align}
      \rho \pdifft{\uu} - \nabla \cdot \bsig &= \mathbf{0}
      \quad \text{in } \Omega,\\
      \nabla \cdot \uu &= 0 
      \quad  \text{in } \Omega,\label{eq:stokes_eqs_cont} 
  \end{align}
  \label{eq:stokes_eqs}% 
\end{subequations}%
where the stress tensor is defined as $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$
and $\beps(\uu) = \frac{1}{2}\left(\nabla \uu + (\nabla\uu)^T\right)$ is the strain-rate tensor. 
Bold-face characters denote vectors or tensors, and $\mathbf{I}$ is the identity tensor in three dimensions.
Furthermore, $\rho=1000 \ \mathrm{kg/m^3}$ is the CSF density
and $\mu=0.7 \ \mathrm{mPa\cdot s}$ is the dynamic viscosity
of the CSF~\cite{Bloomfield1998EffectsFluid}.
We introduce the tangential traction $\bsigpar$ as the
tangential component of the traction $\hat{\bsig}=\bsig\nn$:
\begin{equation*}
    \bsigpar = P_{\nn}(\hat{\bsig}) = (\mathbf{I} - \nn\otimes\nn)\hat{\bsig},
\end{equation*}
where $\nn$ is the outer unit normal of the boundary surface $\Gamma = \partial\Omega$,
and $P_{\nn}(\rr)$ is the tangential projection of a vector $\rr$ onto $\Gamma$.
We consider three types of boundary conditions:
(i) \emph{tangential traction} (no flow normal to the boundary
$\uu \cdot \nn = 0$ with a stress $\btau$ applied tangentially on the boundary $\bsigpar = \btau$),
(ii) \emph{normal pressure} (a pressure $\Tilde{p}$ applied in
the normal direction: $(\mu\nabla\uu - p\mathbf{I})\nn = \Tilde{p}(t)\nn$),
and (iii) \emph{free slip} (no flow normal to the boundary
$\uu \cdot \nn = 0$ and no tangential traction $\bsigpar = {\bf 0}$).
We assume that the system starts at rest: $\uu(\xx, t=0) = \mathbf{0}$.

To study the impact of the motile cilia and the cardiac pulsations on the CSF flow,
we consider the following computational flow scenarios,
each predicting a CSF flow velocity $\uu$ and pressure $p$.
In all of the scenarios, $\Gs$ denotes a free-slip boundary.
\begin{itemize}
    \item A \emph{baseline} flow model representing a wild-type zebrafish including 
    contributions both from the cilia forces and the cardiac pulsations.
    In this scenario, we let $\Gamma=\Gs\cup\Gc\cup\Gp$ and apply the tangential traction $\btau$ on the cilia boundary
    $\Gc$ and prescribe the normal pressure $\Tilde{p}$ on the anterior/posterior boundary $\Gp$. % Model 0/Model C

    \item A \emph{cilia-driven/no-cardiac} flow model including contributions from the
    cilia forces but no cardiac pulsations.
    Here, $\Gamma=\Gs\cup\Gc$, and we apply a tangential traction $\btau$ on the cilia boundary $\Gc$.
    In the absence of cardiac-induced flow, the anterior/posterior boundary $\Gp=\emptyset$. % Model I/Model A

    \item A \emph{cardiac-induced/no-cilia} flow model including contributions from the
    cardiac pulsations but no cilia forces.
    In this scenario, we disregard the cilia and consider $\Gamma=\Gs\cup\Gp$, prescribing the normal pressure $\Tilde{p}$ on the
    anterior/posterior boundary $\Gp$.  % Model II/Model B
\end{itemize}
The tangential traction and normal pressure are described further below.

\subsubsection{Cilia-driven flow}
On the ciliated regions of the ventricle walls $\Gc$
(\Crefpanels{fig:fig1}{d}, magenta markers), we impose a
constant-in-time tangential traction $\bsigpar=\btau$ to represent the
net forces of the cilia acting on the CSF.  We set $\btau(\xx, \rr) =
\tau \lambda(\xx) P_{\nn}(\rr)$, where the function $\lambda(\xx)$ and
the sign of the vector $\rr=\pm(1, 0, 1)$ varies depending on the
cilia region (see~\Cref{subsec:appendixA1}). Previous
experiments~\cite{Olstad2019CiliaryDevelopment} indicate flow speeds
proximate to the ventricular walls of $27.4 \pm 5.4$ \textmu m/s and
$2.6 \pm 0.6$ \textmu m/s in the dorsal and ventral regions of the
middle ventricle, respectively.  In the anterior ventricle, the speeds
were estimated at $4.7 \pm 1.3$ \textmu m/s. We calibrated $\rr$ and
$\tau=0.65$ mPa by comparison of the simulated maximum velocity
magnitude with these flow speeds.

\subsubsection{Flow induced by cardiac pulsations}
To generate a cardiac-induced pulsatile flow inside the brain ventricles, we set $\Tilde{p}(t)=0$ and
$\Tilde{p}(t)=-A\sin{\omega t}$ on anterior and posterior parts, respectively, of the
boundary $\Gp$~(\Crefpanels{fig:fig1}{d}, blue markers).
Based on a measured cardiac frequency of $f=2.22$ Hz~\cite{Olstad2019CiliaryDevelopment},
we set the angular cardiac frequency $\omega=2\pi f=6.97$ rad/s. The amplitude $A=1.5$ mPa determines the magnitude of the pulsatile flow
and was calibrated based on previous experiments~\cite{Olstad2019CiliaryDevelopment}.
Particle tracking data were used to determine the contributions of the
cardiac pulsations to the velocity magnitude,
as compared to the effect of cilia motion on the flow (cf.~\Cref{sec:appC_particle_transport}).

\subsection{Computational model of solute transport after photoconversion}
We simulate transport of photoconverted proteins within the brain ventricles by
modeling transport of a solute with concentration $c(\xx, t)$ for $\xx\in\Omega$ and
time $t > 0$ by the advection-diffusion equation
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0 \quad \text{in } \Omega\setminus\Omega_{\mathrm{pc}},
    \label{eq:adv_diff_strong}
\end{equation}
where the total concentration flux $\JJ$ is assumed to consist of an advective and a diffusive flux:
\begin{equation*}
    \JJ = c\uu - D\nabla c .
\end{equation*}
Here, $\uu$ is the CSF velocity field, governed by the Stokes equations (\Cref{eq:stokes_eqs}),
and $D$ is the molecular diffusion coefficient of the solute. 
The ventricular walls are assumed to be impermeable (admitting \emph{no flux}: $\JJ\cdot\nn=0$),
except at the anterior/posterior boundary $\Gp$.
At the anterior/posterior boundary, we permit the solute concentration to be advected into
and out of the domain in a manner that ensures conservation of mass
(see~\Cref{subsec:adv_diff_eq_weak_form} for details).

To computationally represent protein photoconversion in a region
$\Omega_{\mathrm{pc}} \subset \Omega$ (corresponding to ROI 1),
we prescribe a given time-dependent value to the solute concentration in this region. 
Mimicking the experimental protocol, we consider the dorsal part of the
diencephalic ventricle as the subdomain $\Omega_{\mathrm{pc}}$, and set
\begin{equation}
  c(\xx, t)=\frac{\log{(1+t/a)}}{\log{(1+T/a)}},
  \quad \xx \in \Omega_{\mathrm{pc}}, \quad t > 0.
\label{eq:photoconversion_curve}
\end{equation}
The value $a=65$ was chosen to fit the fluorescence intensity curve
observed in the physical photoconversion experiments. Because steep gradients
in the concentration field may arise from imposing $c(\xx, t)$ in $\Omega_{\mathrm{pc}}$,
the mesh is locally refined around this region.

In order to impose a smooth initial condition, we initially solve the diffusion equation
\begin{equation*}
    \pdifft{c} = \Tilde{D}\nabla^2 c
\end{equation*}
with an increased diffusion coefficient $\Tilde{D}=D\times 10^{5}$ for two timesteps.
The solutions are normalized by their respective maximum values, and subsequently
multiplied with the value of \Cref{eq:photoconversion_curve} at the times $t=\Delta t$
and $t=2\Delta t$, respectively.
The resulting concentration fields are used as initial conditions for $c(\xx, 0)$ and
$c(\xx, \Delta t)$ for all $\xx\in\Omega$, where $\Delta t$
is the timestep. The two initial conditions are required because of the
two-step temporal discretization scheme employed (see~\Cref{subsec:num_approx_adv_diff} for more details.)

\subsection{Estimation of diffusion coefficients via the Stokes-Einstein relation}
Motivated by an interest in molecules that are significant in neural development,
we study the transport of molecules resembling (i) the Dendra2 Fluorescent Protein (Dendra2)
used in the photoconversion, (ii) the Starmaker+Green Fluorescent Protein (STM-GFP) reported by
Jeong~\emph{et al.}~\cite{Jeong2024TheZebrafish}, and (iii) extracellular vesicles.
To estimate the diffusion coefficients of these, when not
available in the literature, we use the Stokes--Einstein relation
\begin{equation}
    D = \frac{k_B T}{6\pi \mu R},
    \label{eq:D_stokes_einstein}
\end{equation}
which is an expression for the diffusion coefficient of a spherical
molecule suspended in a viscous fluid~\cite{Einstein1905UberTeilchen}.
In~\Cref{eq:D_stokes_einstein}, $k_B = \num{1.38e-23}$ J/K is the Boltzmann constant,
$T = 310$ K is the absolute temperature, $\mu$ is the dynamic viscosity of the
CSF, and $R$ is the radius of the molecule.
For the extracellular vesicles, we calculate $D$ using~\Cref{eq:D_stokes_einstein}
with the relatively large $R=150$ nm~\cite{Moghassemi2024ExtracellularDecade}.
For STM-GFP, we estimate the diffusion coefficient by extrapolating
the experimentally observed diffusion coefficient of
GFP~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells},
assuming that if $D$ scales with $1/R$ according to~\Cref{eq:D_stokes_einstein},
$D$ would scale with the cube root of the molecule mass~\cite{Goodhill1997DiffusionGuidance}.
The resulting diffusion coefficients
are $D_1$ (extracellular vesicles), $D_2$ (STM-GFP) and $D_3$ (Dendra2) as given by \Cref{tab:diff_coeff_table}.

\begin{table}[h!]
    \centering
    \caption{Molecular mass and diffusion coefficients of simulated solutes.
    STM and GFP denote Starmaker Protein and Green Fluorescent Protein, respectively.
    (*) The value was extrapolated based on the value of $D$ for
    GFP reported in the literature~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells},
    assuming that $D$ scales with the cube root of the molecule mass.
    (\textdagger) The value was estimated using the Stokes-Einstein relation~\Cref{eq:D_stokes_einstein}.}
    \label{tab:diff_coeff_table}
    \begin{tabular}{lc|ccc}
        \toprule
        Molecule & Label & Mass [kDa]  & $D \ [\mathrm{m^2/s}]$ & Reference \\ 
        \midrule 
        Extracellular vesicles (radius of 150 nm) & $D_1$ & -- & $\num{2.17e-12}$ & \textdagger \\
        GFP (saline aqueous solution) & -- & 26.9~\cite{UniProt2024GFPUniProtKB} & $\num{8.70e-11}$ & \cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells} \\
        STM-GFP & $D_2$ & 66.2~\cite{UniProt2024StmUniProtKB} + 26.9~\cite{UniProt2024GFPUniProtKB} & $\num{5.75e-11}$ & * \\
        Dendra2 Fluorescent Protein & $D_3$ & 25.6~\cite{Gurskaya2006EngineeringLight} & $\num{1.15e-10}$ & \cite{GuraSadovsky2017MeasurementExpansion}\\
        \bottomrule
    \end{tabular}
\end{table}

\subsection{Quantities of interest}
We calculate the mean solute concentrations
\begin{equation}
    \cbar_i(t) = \frac{1}{|\Omega_i|}\int_{\Omega_i} c(\xx, t)\,\dx,
    \label{eq:c_mean_i}
\end{equation}
in the six regions of interest (ROIs) $\Omega_i$ ($i = 1, \dots, 6$)
as functions of time, where $|\Omega_i|$ is the volume of
the domain $\Omega_i$. The dynamics of ROI 1 follows directly
from~\Cref{eq:photoconversion_curve}. In addition,
we report the time-to-threshold as the time $\hat{t}_i$ when
the mean concentration $\cbar_i$ first exceeds a threshold
value $\hat{c}_i$. We use $\hat{c}_i=0.25$ for ROI 1--4 and
$\hat{c}_i=0.10$ for ROI 5--6.

To evaluate the relative importance of advection and
diffusion as transport mechanisms, we consider the Péclet number
\begin{equation*}
    \mathrm{Pe} = \frac{\mathrm{diffusion \ timescale}}{\mathrm{advection \ timescale}}
    = \frac{t_d}{t_a} = \frac{L_p^2/D}{L_p/U_p} = \frac{L_p U_p}{D}.
\end{equation*}
Thus, if $\mathrm{Pe} > 1$, transport is dominated by advection,
and if $\mathrm{Pe} < 1$ transport is dominated by diffusion.
Here, $L_p$ and $U_p$ are characteristic length and velocity scales,
respectively, and $D$ is the solute diffusion coefficient.
In calculating a global Péclet number, we use $L_p=600$ \textmu m, the approximate length of the
ventricular geometry along the rostrocaudal ($x$-) axis.
We use a mean velocity $\overline{U}$ as the characteristic velocity $U_p$.
The mean velocity $\overline{U}$ is calculated by first averaging the velocity in space:
\begin{equation*}
    \overline{u}^2 = \frac{1}{|\Omega|}\int_{\Omega}\uu\cdot\uu\dx,
\end{equation*}
and then averaging $\overline{u}$ in time over one cardiac cycle to yield $\overline{U} = 2.4$ \textmu m/s.
To assess the assumption that inertial forces are negligible compared to
viscous forces, we also compute the Reynolds number
\begin{equation*}
    \mathrm{Re} = \frac{\mathrm{inertial \ forces}}{\mathrm{viscous \ forces}} = \frac{\rho U_r L_r}{\mu},
\end{equation*}
with $L_r = 110$ \textmu m (the height of the middle ventricle) and
$U_r = 27.9$ \textmu m/s (the maximum velocity magnitude)
as length and velocity scales characteristic of the flow.

To study how cilia paralysis and ventricular morphology affect intraventricular
flow and transport, we calculate the magnitude of a tangential force $F$ 
exerted by the cilia on a part $\mathcal{S}\subset\Gc$ of the ventricular surface as:
\begin{equation}
    F = \int_{\mathcal{S}}\sqrt{\btau\cdot\btau}\,\mathrm ds.
    \label{eq:tangential_force_calculation}
\end{equation}

\subsection{Numerical approximation of the Stokes equations}
The Stokes equations (\Cref{eq:stokes_eqs}) are discretized with a finite element method
in space, while in time, we employ a first-order implicit Euler discretization~\cite{butcher2016numerical}.
We set the timestep size $\Delta t = 0.023$ s such that 20 timesteps represent one cardiac cycle.
For the spatial discretization, we use piecewise linear Brezzi-Douglas-Marini (BDM)
elements for the velocity.~\cite{Brezzi1985TwoProblems, Nedelec1986AIR3}
The BDM elements have degrees of freedom associated with integral moments of the
velocity component normal to the facets of elements,
thus we enforce the impermeability condition $\uu\cdot\nn=0$ strongly.
We impose the boundary conditions for both the tangential traction $\bsigpar$ and
the normal pressure $\Tilde{p}(t)$ weakly.
For the pressure, we use piecewise constant (zeroth order discontinuous Galerkin (DG)) elements.
The resulting total number of degrees of freedom at each time step is 1 009 874.
The element combination $\mathrm{BDM}_1$--$\mathrm{DG}_{0}$ yields a non-conforming
discretization scheme for the Stokes equations~\cite{Stenberg1989SomeEquations}.
Continuity in tangential velocities on interior facets
is weakly enforced with a penalty parameter $\gamma=10$~\cite{Hong2016AEquations}.
Since the divergence of the velocity space is a subspace of the pressure space for this scheme,
the divergence-free condition of the velocity is satisfied exactly, ensuring mass conservation.~\cite{Boffi2008FiniteProblem}

\subsection{Numerical approximation of the advection-diffusion equation}\label{subsec:num_approx_adv_diff}
The advection-diffusion equation (\Cref{eq:adv_diff_strong}) is discretized
with with a second-order, symmetric interior penalty DG
method in space~\cite{Arnold1982AnElements}, chosen due to its favorable
conservation properties. At each timestep the total number of degrees of freedom 
is 1 415 120. In time, the equation is discretized with the second-order backward
difference scheme BDF2.~\cite{Volker2016}
To accurately approximate the solute concentrations even
in the presence of strong advection (high Péclet numbers),
we use an upwind scheme for the velocity when calculating the
advective flux~\cite{Patankar2018NumericalFlow}.
As for the flow equations, we use a timestep size of
$\Delta t = 0.023$ s. We simulate transport for 1900 cardiac cycles,
with a final time of simulation $T = 856$ s.
A DG penalty parameter $\alpha = 50$ was chosen to ensure stability of the method.

\subsection{Solution strategy and software implementation}
Owing to the one-way coupling between the velocity field $\uu$
and the concentration $c$, the governing equations for the
CSF flow and the concentration can be solved sequentially.
For the cilia-driven/no-cardiac flow scenario, the velocity quickly reaches steady-state since
the viscous forces dominate the inertial forces.
In this case, we first solve the steady-state Stokes equations (\Cref{eq:stokes_eqs})
for $\uu$ and $p$, and then use the velocity field $\uu$ as input when solving
the advection-diffusion equation (\Cref{eq:adv_diff_strong}).
For the other flow scenarios, the Stokes equations are solved for one
period of the pulsatile cardiac motion. The solution of this one period
is then used periodically as input for the advection-diffusion equation.

The numerical solution schemes for the Stokes equations and the
advection-diffusion equation were implemented and solved numerically with
DOLFINx with petsc4py as linear algebra backend
\cite{dolfinx2023preprint, Dalcin2011ParallelPython}.
The simulation code is openly
available\fixme{~\cite{zenodo-link-to-all-code}}.
A direct solver using MUMPS~\cite{Amestoy2011Mumps} is used to solve the
linear systems resulting from the discretized Stokes equations.
For the discretized advection-diffusion equation, we use a flexible GMRES solver
with block Jacobi preconditioning \cite{fgmres, Jacobi1845UeberGleichungen}. 
We discretize the time-dependent diffusion equation (without advection)
and use the assembled linear system matrix
as the preconditioner matrix for the advection-diffusion problem.
Numerical experiments verified the accuracy of the model implementation
(\Cref{sec:appB_verification}).

% Acknowledgements
\medskip
\noindent \textbf{Acknowledgments} \par
\noindent Thanks are extended to J\o rgen Dokken
for valuable model implementation discussions.  H.~H.~and
M.~E.~R. acknowledge support from the national infrastructure for
computational science in Norway, Sigma2, via grant
\#NN8049K. M.~E.~R.~acknowledges support from Stiftelsen Kristian
Gerhard Jebsen via the K.~G.~Jebsen Centre for Brain Fluid Research
and support from the Research Council of Norway (RCN) via grant
\#324239 (EMIx). N.~J.~Y.~acknowledges support from the Research
Council of Norway (grants \#314189 and \#353348).

% Data Availability Statement
\medskip
\noindent \textbf{Data Availability Statement} \par
\noindent The data and implementations presented in this paper are licensed under \fixme{LICENSE}
and openly available at \fixme{REPO-LINK}.

% Conflic of Interest Disclosure
\medskip
\noindent \textbf{Conflict of Interest Disclosure} \par
\noindent The authors declare no conflicts of interest.

% Bibliography
\bibliographystyle{MSP}
\bibliography{ref}

\newpage 
%%--------------APPENDIX--------------%%
\appendix
\section{Supplementary Theory}
\subsection{Weak formulation of the Stokes equations}\label{subsec:appendixA1}
We consider the time-dependent Stokes equations
\begin{subequations}
\begin{alignat}{2}
   \rho\pdifft{\uu} - \nabla \cdot \bsig(\uu, p)
   &= \mathbf{0} \quad &&\mathrm{in} \ \Omega, \label{eq:BDM_stokes_mom}\\
  \nabla \cdot \uu &= 0 \quad &&\mathrm{in} \ \Omega, \label{eq:BDM_stokes_div} 
\end{alignat}
\label{eq:appendix_stokes_eqs}% 
\end{subequations}%
in the domain $\Omega$ with boundary $\Gamma$, where
$\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$ is the viscous
stress tensor, in which $\beps(\uu) = \frac{1}{2}(\nabla\uu + (\nabla\uu)^T)$
is the symmetric gradient. Furthermore, $\uu\in\VV$ is the fluid velocity
and $p\in Q$ is the fluid pressure, and $\rho$ and $\mu$ are the fluid density
and dynamic viscosity, respectively. Unless otherwise stated, we let $\bsig = \bsig(\uu, p)$.

To derive weak formulations of the continuous problem defined by the Stokes
equations (\Cref{eq:appendix_stokes_eqs}), we consider the function space
\begin{align*}
    \VV &= \bigg\{\vv \ : \ \vv\in\mathbf{H}^1(\Omega),
    \vv\cdot\nn=0 \ \mathrm{on} \ \Gamma\setminus\Gp\bigg\},
\end{align*}
for the velocity and $Q=L^2(\Omega)$ for the pressure.
Here $\Gp$ is the anterior/posterior boundary where we impose a normal pressure
boundary condition for the model versions that consider fluid motion induced by
cardiac pulsations. Bold-faced characters for a space denotes a $d$--dimensional vector space,
where $d$ is the spatial dimension of $\Omega$. For example,
$\mathbf{H}^1(\Omega) = [H^1(\Omega)]^3$ if $\Omega$ is three-dimensional.

To derive a weak formulation of the Stokes equations (\Cref{eq:appendix_stokes_eqs}),
we begin by multiplying~\Cref{eq:BDM_stokes_mom} with a
test function $\vv\in\VV$ and integrating over the domain:
\begin{equation*}
    \intO{\rho\pdifft{\uu}\cdot\vv} - \intO{(\nabla\cdot\bsig)\cdot\vv} = \mathbf{0}.
\end{equation*}
Performing integration by parts of the stress term and applying Green's first identity, we get
\begin{equation}
    \intO{\rho\pdifft{\uu}\cdot\vv}
    - \intG{\bsig\nn\cdot\vv}
    + \intO{2\mu\beps(\uu) : \beps(\vv)}
    - \intO{p\nabla\cdot\vv} = \mathbf{0}, \label{eq:modelA_weak_form1}
\end{equation}
where the symbol : signifies the matrix inner product,
$A:B=\text{trace}{(A^TB)}$ for matrices $A, B\in\mathbb{R}^{d\times d}$.

Further steps depend on the boundary conditions applied,
which is different for the three model versions we consider in this paper.
Before proceeding, let us introduce some notation used in defining the
slip boundary conditions. We make use of a decomposition
of any vector $\vv$ on the boundary $\Gamma$ into its normal and tangential components,
$\vv = \vv_{\perp} + \vv_{\parallel}$, where $\vv_{\perp}=(\vv\cdot\nn)\nn$ and
$\vv_{\parallel}=P_{\nn}(\vv)$. The tangential projection is defined as
\begin{equation}
    P_{\nn}(\vv) = (\mathbf{I} - \nn\otimes\nn)\vv,
    \label{eq:projection_operator_appendix}
\end{equation}
where $\nn$ is the outer unit normal of the surface. Introducing
$\hat{\bsig} = \bsig\nn$ for the traction (vector), the forcing due
to cilia is represented by constraining $\bsigpar$ on the part of the
boundary populated by cilia. In particular, we let
$\bsigpar = \btau(\xx, \rr) = \tau\lambda(\xx) P_{\nn}(\rr)$,
where the parameter $\tau=0.65$ mPa was used to control the magnitude
of the tangential stresses, and chosen by calibrating simulated velocity
fields with experimental data~\cite{Olstad2019CiliaryDevelopment}.
The details of the function $\lambda(\xx)$ follow subsequent to deriving the weak form.

We choose $\rr = \rr^+ = (1, 0, 1)$ or $\rr = \rr^- = -(1, 0, 1)$
to impose forces perpendicular to the $y$-axis.
The reason for this is that all of the experimental images we compare
with are taken in the $xz$-plane (rostrocaudal-dorsoventral plane),
rendering flow features in the $y$-direction (lateral) unobservable.
In the following, we consider the way of handling the stress
boundary integral for each model version successively.

\textbf{Cilia-driven/no-cardiac model.} This model version has cilia forces
as the only mechanism driving flow. We consider the
Stokes equations~\Cref{eq:appendix_stokes_eqs} with the slip boundary conditions
\begin{subequations}
    \begin{alignat}{2}
        \uu\cdot\nn &= 0 \quad &&\mathrm{on} \ \Gamma=\Gc\cup\Gs,
        \label{eq:modelA_slip_BC_impermeability} \\
        \bsigpar &= \btau \quad &&\mathrm{on} \ \Gc,
        \label{eq:modelA_slip_BC_shear} \\
        \bsigpar &= \mathbf{0} \quad &&\mathrm{on} \ \Gs,\label{eq:modelA_full_slip_BC}
    \end{alignat}\label{eq:modelA_slip_BC}%
\end{subequations}%
on the boundary $\Gamma=\Gc\cup\Gs$. The boundary $\Gs$ is a free-slip boundary,
where we impose the free-slip condition given by \Cref{eq:modelA_full_slip_BC}.
The first condition (\Cref{eq:modelA_slip_BC_impermeability}) is the
impermeability condition, enforcing no flow normal to the boundary,
whereas~\Cref{eq:modelA_slip_BC_shear} imposes a traction tangential to
the boundary. This tangential traction is applied to represent
the forces that the collective cilia motion exerts on the cerebrospinal fluid.

The boundary conditions defined by~\Cref{eq:modelA_slip_BC}
are set weakly through the boundary integral 
\begin{equation}
    \intG{\bsig\nn\cdot\vv} =  \intG{\hat{\bsig}\cdot\vv}
    \label{eq:boundary_integral_BC1}
\end{equation}
in the weak form of the Stokes equations (\Cref{eq:modelA_weak_form1}).
We decompose the traction vector $\hat{\bsig}$ into a normal and tangential component:
\begin{equation*}
    \hat{\bsig} = \bsigperp + \bsigpar.
\end{equation*}
Inserting the above into~\Cref{eq:boundary_integral_BC1},
combined with application of the tangential traction
condition (\Cref{eq:modelA_slip_BC_shear}) and the
free-slip condition (\Cref{eq:modelA_full_slip_BC}), yields
\begin{equation}
    \intG{\hat{\bsig}\cdot\vv} = \int_{\Gc\cup\Gs}{(\bsigperp + \bsigpar)\cdot\vv}\, ds = 
    \int_{\Gc\cup\Gs}{\bsigperp\cdot\vv}\, ds
    + \intGc{\btau\cdot\vv} = \intG{\bsigperp\cdot\vv} + \intGc{\btau\cdot\vv}.
    \label{eq:boundary_integral_BC2}
\end{equation}
Since $\btau$ is known, the second term on the right-hand side of~\Cref{eq:boundary_integral_BC2}
is moved to the right-hand side of the weak form.
With $\hat{\bsig}_{\perp} = (\hat{\bsig}\cdot\nn)\nn$,
we can write the first term on the right-hand side of~\Cref{eq:boundary_integral_BC2} as
\begin{equation}
    \intG{\bsigperp\cdot\vv} = \intG{((\hat{\bsig}\cdot\nn)\nn)\cdot\vv} =
    \intG{(\hat{\bsig}\cdot\nn)(\vv\cdot\nn)}.
    \label{eq:boundary_integral_BC_normal}
\end{equation}
This term vanishes owing to the construction $\VV$.
As a result, for the cilia-driven/no-cardiac model,
the weak form of the momentum equation takes the following form:
\begin{equation*}
    \intO{\rho\pdifft{\uu}\cdot\vv}
    + \intO{2\mu\beps(\uu) : \beps(\vv)}
    - \intO{p\nabla\cdot\vv} = \intGc{\btau\cdot\vv}, \quad\forall\vv\in\VV.
\end{equation*}
Next, multiplying the continuity equation (\Cref{eq:BDM_stokes_div})
with a test function $q\in Q$, we require that
\begin{equation}
    -\intO{q\nabla\cdot\uu} = 0,\quad\forall q\in Q.
    \label{eq:pressure_weak_form}
\end{equation}
The full weak form of the Stokes equations for this flow model is then:
find $(\uu, \, p) \in \VV\times Q$, such that
\begin{subequations}
    \begin{alignat}{5}
        &\intO{\rho\pdifft{\uu}\cdot\vv}
        + &\intO{2\mu\beps(\uu) : \beps(\vv)}
        - &\intO{p\nabla\cdot\vv}
        &&= \intGc{\btau\cdot\vv}, &&\quad\forall\vv\in\VV, \label{eq:stokes_weak_mom_modelA}\\ 
        &{} &{}  -&\intO{q\nabla\cdot\uu} &&= 0, &&\quad\forall q\in Q.
        \label{eq:stokes_weak_cont_modelA}
    \end{alignat}%
    \label{eq:stokes_weak_modelA}%
\end{subequations}%
We observe that, in this case with slip boundary conditions on the entirety of
the boundary $\Gamma$, for any real constant $C$ the vector
$(\uu, p)=(\mathbf{0}, C)$ satisfies the equations.
Hence \Cref{eq:stokes_weak_modelA} is a singular system 
with a one-dimensional nullspace of constant pressures.
Consequently, assembly of the discretization of the Stokes equations (\Cref{eq:stokes_weak_modelA})
into a linear equation system $\mathbf{A}\mathbf{x}=\mathbf{b}$ yields a singular matrix $\mathbf{A}$.
To ensure existence of solutions, we orthogonalize the right-hand side vector $\mathbf{b}$ with
respect to the pressure nullspace basis (the set of all constants), before solving the linear
system. To obtain a unique solution, which together with existence renders the
problem well-posed, information of the pressure nullspace is passed to the linear solver.
Let us remark that the singularity is a consequence of the fact that the
impermeability condition is assumed to hold on the \emph{entire} domain boundary
for this model version, which leaves the pressure under-determined as a result of no
boundary condition specifying the pressure.

For the tangential traction $\btau = \tau\lambda(\xx) P_{\nn}(\rr)$ that represents the cilia
motion, we heuristically determined the constant $\tau= 0.65$ mPa
and the spatially varying function $\lambda(\xx)$. The definitions of $\rr$ and $\lambda(\xx)$
for the different cilia populations are summarized in~\Cref{tab:r_vector_and_lambda_function}.
\begingroup
\setlength{\tabcolsep}{20pt}
\renewcommand{\arraystretch}{1.5}
\begin{table}[H]
    \centering
    \caption{Definitions of $\rr$ and $\lambda(\xx)$ used in the cilia tangential 
    traction boundary condition $\btau = \tau\lambda(\xx)P_{\nn}(\rr)$,
    where $\tau=0.65$ mPa. The $x-$coordinates $x_{0, i}$ and $x_{e, i}$ define 
    the cilia domain in the dorsal ($i=d$) and ventral ($i=v$) regions of 
    the middle ventricle.}\label{tab:r_vector_and_lambda_function}
    \begin{tabular}{lcc}
        \toprule
        Cilia population & $\rr$ & $\lambda(\xx)$ \\
        \midrule 
        Anterior region, anterior ventricle 
        & $-(1, 0, 1)$ & $\frac{2}{5}$ \\
        Posterior region, anterior ventricle
        & $\ \ \ (1, 0, 1)$ & 1 \\
        Dorsal region, middle ventricle     
        & $-(1, 0, 1)$
        & $\frac{11}{4}\left(\frac{x-x_{0, d}}{x_{e, d} - x_{0, d}}\right)$\\
        Ventral region, middle ventricle    
        & $\ \ \ (1, 0, 1)$
        & $\frac{1}{2}\left(1 - \frac{x-x_{0, v}}{x_{e, v} - x_{0, v}}\right)$ \\
        \bottomrule
    \end{tabular}
\end{table}
\endgroup

\textbf{Cardiac-induced/no-cilia model.} For this model version,
the only physical mechanism inducing fluid flow is the cardiac cycle.
The cardiac cycle generates directional back-and-forth motion of the CSF
in the zebrafish brain ventricles.
We model the influence of the cardiac cycle with a normal pressure boundary condition:
\begin{subequations}
    \begin{alignat}{2}
      \uu\cdot\nn = 0, \ \hat{\bsig}_{\parallel} &= \mathbf{0} \quad &&\mathrm{on} \ \Gs,
      \label{eq:modelB_slip_BC} \\
      (\mu\nabla\uu - p\mathbf{I})\nn &= \Tilde{p}(t)\nn \quad &&\mathrm{on} \ \Gp,
      \label{eq:modelB_pressure_BC}
    \end{alignat}%
\end{subequations}%
where the boundary $\Gamma = \Gs\cup\Gp$ consists of a free-slip boundary $\Gs$,
and a part $\Gp$ where a normal traction type of boundary condition is imposed.
Note, however, that the transpose velocity gradient term must be
included inside the parentheses on the left-hand side of~\Cref{eq:modelB_pressure_BC}
for the expression to be the stress tensor $\bsig$, thus the left-hand side of
\Cref{eq:modelB_pressure_BC} does not constitute the normal traction $\bsigperp$.
Because of this, we rather refer to this boundary condition
as a normal pressure boundary condition.
The boundary $\Gs$ is handled as discussed in the previous section,
such that the stress boundary integral in~\Cref{eq:boundary_integral_BC1}
vanishes over $\Gs$. Therefore, we now have
\begin{align}
    \intG{\hat{\bsig}\cdot\vv} = \intGp{\hat{\bsig}\cdot\vv}
    &= \intGp{(\mu(\nabla\uu + (\nabla\uu)^T) - p\mathbf{I})\nn\cdot\vv}\notag \\
    &= \intGp{\mu(\nabla\uu)^T\nn\cdot\vv} + \intGp{(\mu\nabla\uu
    - p\mathbf{I})\nn\cdot\vv}\notag\\
    &= \intGp{\mu(\nabla\uu)^T\nn\cdot\vv} + \intGp{\Tilde{p}(t)(\vv\cdot\nn)}.
    \label{eq:modelB_boundary_integral1}
\end{align}
We leave the first term on the left-hand side
of~\Cref{eq:modelB_boundary_integral1} in the weak form of
the momentum equation. The second term is determined by
the boundary condition defined in~\Cref{eq:modelB_pressure_BC},
and is therefore moved to the right-hand side of the weak form.
The weak form of the momentum equation is then
\begin{equation}
    \begin{aligned}
    \intO{\rho\pdifft{\uu}\cdot\vv} + \intO{2\mu\beps(\uu) : \beps(\vv)} -
    \intO{p\nabla\cdot\vv} &-\intGp{\mu(\nabla\uu)^T\nn\cdot\vv} \\
    &=\intGp{\Tilde{p}(t)(\vv\cdot\nn)}, \quad\forall\vv\in\VV.
    \end{aligned}
    \label{eq:weak_form_mom_eq_modelB}
\end{equation}
Also considering the weak form of the continuity equation,
the weak formulation for the cardiac-induced/no-cilia model version
then reads: find $(\uu,\, p)\in\VV\times Q$
such that~\Cref{eq:weak_form_mom_eq_modelB} and~\Cref{eq:pressure_weak_form}
hold for all $(\vv, \, q)\in\VV\times Q$.
To simulate the back-and-forth motion induced by the cardiac cycle observed in
previous work~\cite{Olstad2019CiliaryDevelopment}, we apply sinusoidal forcing.
The boundary $\Gp$ has two separate regions: one located on the
anterior ventricle and one on the posterior ventricle~(\Crefpanels{fig:fig1}{d}, blue markers).
For the normal pressure boundary condition, we set $\Tilde{p}(t) = 0$ on the anterior part and
$\Tilde{p}(t) = p_c(t) = -A\sin{\omega t}$ on the posterior part. The value of the amplitude
$A$ was based on heuristics and $\omega = 2\pi f$, where $f=2.22$ Hz
is the cardiac cycle frequency.

\textbf{Baseline model.} The baseline flow model
considers both cilia and cardiac beating as driving mechanisms of fluid flow.
We therefore consider the boundary conditions
\begin{subequations}
    \begin{alignat}{2}
      \uu\cdot\nn = 0, \ \bsigpar
      &= \btau  \quad &&\mathrm{on} \ \Gc, \label{eq:slip_BC_modelC} \\
      \uu\cdot\nn = 0, \ \bsigpar
      &= \mathbf{0}  \quad &&\mathrm{on} \ \Gs, \label{eq:freeslip_BC_modelC} \\
      (\mu\nabla\uu - p\mathbf{I})\nn
      &= \Tilde{p}(t)\nn \quad &&\mathrm{on} \ \Gp, \label{eq:pressure_BC_modelC}
    \end{alignat}
\end{subequations}%
where $\Gamma = \Gc\cup\Gs\cup\Gp$ is the boundary of $\Omega$.
The quantities $\btau$ and $\Tilde{p}(t)$ have been introduced in
the previous sections covering the other flow model versions. We handle the boundary
conditions~\Cref{eq:slip_BC_modelC} and~\Cref{eq:freeslip_BC_modelC}
as introduced for the cilia-driven/no-cardiac model, and the boundary
condition~\Cref{eq:pressure_BC_modelC} as introduced for the cardiac-induced/no-cilia model.
The abstract weak formulation for 
the baseline model is then: find $(\uu,\, p)\in\VV\times Q$ such that
\begin{subequations}
    \begin{alignat}{2}
        \mathcal{A}(\uu, \vv) + \mathcal{B}(\vv, p) &= \mathcal{L}(\vv),
        &&\quad\forall\vv\in\VV, \\
        \mathcal{B}(\uu, q) &= 0, &&\quad\forall q\in Q,
    \end{alignat}%
    \label{eq:abstract_weak_form_modelC}
\end{subequations}%
with the definitions
\begin{align}
    \mathcal{A}(\uu, \vv) &= \intO{\rho\pdifft{\uu}\cdot\vv}
                          + \intO{2\mu\beps(\uu) : \beps(\vv)}
                          -\intGp{\mu(\nabla\uu)^T\nn\cdot\vv},\label{eq:stokes_bilinear_form}\\
    \mathcal{B}(\uu, q) &= -\intO{q\nabla\cdot\uu}, \label{eq:stokes_skew_form}\\
    \mathcal{L}(\vv) &= \intGc{\btau\cdot\vv} + \intGp{\Tilde{p}(t)(\vv\cdot\nn)}.\label{eq:stokes_linear_form}
\end{align}

\subsection{Discretization of the Stokes equations}
We outline the discretization of the weak
formulation of the Stokes equations for the baseline model 
(both cilia traction and cardiac pressure pulsations present) given by
\Cref{eq:abstract_weak_form_modelC,eq:stokes_bilinear_form,eq:stokes_skew_form,eq:stokes_linear_form}.
A similar approach can be used to discretize the weak formulation
for the other model versions.
First, we consider the spatial discretization of the weak formulation.
We rewrite the weak form in terms of the discretized variables,
denoted with the subscript $h$, on the discrete domain $\Omega_h$:
\begin{subequations}
    \begin{alignat}{2}
        \mathcal{A}(\uu_h, \vv_h) + \mathcal{B}(\vv_h, p_h)
        &= \mathcal{L}(\vv_h), &&\quad\forall\vv_h\in\VV_h, \\
        \mathcal{B}(\uu_h, q_h) &= 0, &&\quad\forall q_h\in Q_h.
    \end{alignat}%
    \label{eq:no_stabilization_abstract_discrete_weak_form_modelC}%
\end{subequations}%
Our choice of discretization are finite elements
of the Brezzi-Douglas-Marini (BDM) family~\cite{Brezzi1985TwoProblems, Nedelec1986AIR3}
for the velocity
space and discontinuous Lagrange polynomial elements for the pressure. We refer to the latter
as discontinuous Galerkin (DG) elements. For all cells $K$ of the computational mesh $\Omega_h$,
we define
\begin{align*}
    \VV_h &= \Big\{\vv_h : \vv_h\in\mathbf{H}(\mathrm{div});
    \vv_h\vert_K\in\big[P_k(K)\big]^d;
    \forall K\in\Omega_h;
    \vv_h\cdot\nn=0 \ \text{on } \Gamma\setminus\Gp\Big\},\\
    Q_h &= \Big\{q_h : q_h\in L^2;  q_h\vert_K\in P_{k-1}(K); \forall K\in\Omega_h\Big\},
\end{align*}
where $P_k$ is the space of polynomials of degree less than or equal to $k$.
The BDM family of elements have degrees of freedom
associated with integral moments of vector components normal to facets of the mesh.
Because $\VV_h\not\subset\VV$, this BDM-DG scheme yields a non-conforming
method for the Stokes equations. The scheme is
$\mathbf{H}(\mathrm{div})$-conforming and exactly mass-conserving:
due to the fact that $\nabla\cdot\VV_h\subset Q_h$,
the continuity equation (\Cref{eq:BDM_stokes_div})
is satisfied in every computational mesh cell.
Moreover, control of normal components of $\uu_h$ conveniently 
lets us enforce the impermeability condition $\uu_h\cdot\nn=0$ strongly.

Because the elements of $\VV_h$ only define the normal components of $\uu_h$,
continuity in tangential components of $\uu_h$ on interior facets of
the mesh is ensured through stabilization
of the discrete weak form (\Cref{eq:no_stabilization_abstract_discrete_weak_form_modelC}).
We apply the stabilization outlined in Hong~\emph{et al.}~\cite{Hong2016AEquations}.
First, we define the average and jump operators, respectively, as
\begin{equation}
    \avg{\uu} = \frac{\uu^+\cdot\nn^+ - \uu^-\cdot\nn^-}{2}
    \qquad\mathrm{and}\qquad \jump{\uu} = \uu^+ - \uu^-,
    \label{eq:operators_BDM}
\end{equation}
where $\uu$ is a first-order tensor and 
$\nn$ is the normal vector on a facet separating two cells,
and the plus and minus signs denote variables in each of the two cells.
The normal vector is chosen such that it points towards the exterior of
its corresponding cell. On the boundary, the operators are defined as
$\avg{\uu} = \uu\cdot\nn$ and $\jump{\uu} = \uu$,
with $\nn$ pointing to the exterior of the domain. Note that average operator
defined in~\Cref{eq:operators_BDM} reduces the rank of the tensor $\uu$ by one,
whereas the jump operator preserves rank. This extends to tensors of higher order.
We now define and add the bilinear form
\begin{align*}
    \mathcal{A}_{\mathrm{stab}}(\uu_h, \vv_h) &=
    \sum_{f\in \mathcal{F}}\int_f{2\mu\frac{\gamma}{\avg{h}}\jump{P_{\nn}(\uu_h)}
    \cdot\jump{P_{\nn}(\vv_h)}}\,\mathrm ds\\
    &-\sum_{f\in \mathcal{F}}\int_f{2\mu\avg{\beps(\uu_h)}\cdot\jump{\vv_h}}\,\mathrm ds
    -\sum_{f\in \mathcal{F}}\int_f{2\mu\avg{\beps(\vv_h)}\cdot\jump{\uu_h}}\,\mathrm ds
\end{align*}
to the discrete weak formulation (\Cref{eq:no_stabilization_abstract_discrete_weak_form_modelC}),
where $\mathcal{F}$ is the set of all interior facets of the computational mesh.
The first term of $\mathcal{A}_{\mathrm{stab}}$ penalizes jumps
in the tangential components of the velocity
across interior facets, where the penalty parameter $\gamma>0$ must be chosen large enough
to achieve stability. For appropriate convergence, the penalization term
includes a scaling with the average of the cell diameter of neighboring mesh cells,
with $h$ defined as the maximal distance between two points in a cell~\cite{Nitsche1971UberSind}.
The other two terms
of $\mathcal{A}_{\mathrm{stab}}$ are added to ensure consistency and symmetry of
the bilinear form, such that the problem is well-defined.

Including the stabilization outlined above, the semi-discrete weak formulation of
the Stokes equations reads:
find ($\uu_h, p_h$)$\in$($\VV_h, Q_h$) such that
\begin{subequations}
    \begin{alignat}{2}
        \mathcal{A}(\uu_h, \vv_h)
        + \mathcal{A}_{\mathrm{stab}}(\uu_h, \vv_h)
        + \mathcal{B}(\vv_h, p_h)
        &= \mathcal{L}(\vv_h), &&\quad\forall\vv_h\in\VV_h, \\
        \mathcal{B}(\uu_h, q_h) &= 0, &&\quad\forall q_h\in Q_h.
    \end{alignat}%
    \label{eq:abstract_semi_discrete_weak_form_modelC}%
\end{subequations}%
In time, we discretize the weak formulation (\Cref{eq:abstract_semi_discrete_weak_form_modelC})
with the implicit Euler method, approximating the time derivative as
\begin{equation*}
    \pdifft{\uu} \approx \frac{\uu^{n+1} - \uu^n}{\Delta t},
\end{equation*}
where $n$ denotes the timestep index. All of the other terms
in~\Cref{eq:abstract_semi_discrete_weak_form_modelC} are evaluated at timestep $n+1$.
The method is implicit and of first order in time~\cite{butcher2016numerical}. 
The fully discrete Stokes equations were solved with the penalty parameter $\gamma=10$.

\subsection{Weak formulation of the advection-diffusion equation}
\label{subsec:adv_diff_eq_weak_form}
Recall the strong form of the advection-diffusion equation:
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0 \quad\mathrm{in} \ \Omega,
    \label{eq:adv_diff_strong_DG}
\end{equation}
where $\Omega$ is the domain with boundary $\Gamma$, and the total solute flux
\begin{equation*}
    \JJ(c) = c\uu - D\nabla c,
\end{equation*}
where $c\in W=H^1$ is the solute concentration and
$\uu\in\VV$ is the cerebrospinal fluid velocity .
We consider $\JJ = \JJ(c)$, unless otherwise indicated.
The weak form of~\Cref{eq:adv_diff_strong_DG} is derived by
multiplying the equation with a test function $\phi \in W$ and integrating over the domain: 
\begin{equation*}
    \intO{\pdifft{c}\phi} + \intO{(\nabla\cdot\JJ)\phi} = 0.
\end{equation*}
Integrating the second term by parts and applying Green's first identity yields
\begin{equation}
    \intO{\pdifft{c}\phi} + \intG{(\JJ\cdot\nn)\phi}
    - \intO{\JJ\cdot\nabla\phi} = 0.\label{eq:adv_diff_weak_formA1}
\end{equation}
The boundary integral of the total flux is handled differently for
the three different model versions, depending on the boundary conditions considered.

\textbf{Cilia-driven/no-cardiac model.} We consider $\Omega$ to be
a closed domain with impermeable walls, so that the whole boundary
$\Gamma$ constitutes a no-flux boundary $\Gnf$.
By imposing $\JJ\cdot\nn=0$ on $\Gamma=\Gnf$, 
the weak form defined by \Cref{eq:adv_diff_weak_formA1} is reduced to
\begin{equation*}
    \intO{\pdifft{c}\phi} - \intO{\JJ\cdot\nabla\phi} = 0, \quad\forall\phi\in W.
\end{equation*}
Defining the left-hand side as the bilinear form
\begin{equation*}
    a(c, \phi) = \intO{\pdifft{c}\phi} - \intO{\JJ\cdot\nabla\phi}, 
\end{equation*}
we get the following abstract weak formulation: find $c\in W$ such that
\begin{equation*}
    a(c, \phi) = 0, \quad\forall\phi\in W.
\end{equation*}

\textbf{Cardiac-induced/no-cilia model.} When using the normal pressure
boundary condition (\Cref{eq:modelB_pressure_BC}) to model pulsatile flow,
we have in- and outflow of cerebrospinal fluid and thus advection of
$c$ on the pressure boundary $\Gp$.
Let $\Gp=\Gin\cup\Gout$ and $\Gamma = \Gnf\cup\Gin\cup\Gout$, with 
\begin{align*}
    \Gin  &= \Big\{\xx\in\Gamma : \uu(\xx, t)\cdot\nn(\xx) < 0 \Big\}, \\
    \Gout &= \Big\{\xx\in\Gamma : \uu(\xx, t)\cdot\nn(\xx) > 0 \Big\},
\end{align*}
as in- and outflow boundaries, respectively, where we have emphasized that
both $\uu$ and $\nn$ are spatially dependent variables.
With $\JJ\cdot\nn=0$ on the no-flux boundary $\Gnf$, we have
\begin{equation*}
    \intG{(\JJ\cdot\nn)\phi} = \intGin{(\JJ\cdot\nn)\phi} + \intGout{(\JJ\cdot\nn)\phi}. 
\end{equation*}
For the outflow boundary integral over $\Gout$, we impose zero
diffusive flux $D\nabla c\cdot\nn=0$,
with the advection term $c(\uu\cdot\nn)$ simply left in the variational form as an unknown.
For the continuous problem this is problematic, since we are lacking a boundary
condition to close the problem. 
In the discrete setting, however, this method of leaving the boundary term has been shown
to yield consistent results~\cite{Papanastasiou1992ACondition, Lynch2020NumericalHemodynamics,
Griffiths1997TheCondition}. Intuitively, one can think of this outflow boundary condition as
setting the advective flux to being the one consistent with the weak form,
for which there is a unique $c$ satisfying the equations, with $\uu$ determined by
the Stokes equations.

The flux integral over the inflow boundary $\Gin$ cannot be handled
in the same way as the method used for $\Gout$, as this would not preserve
coercivity of the weak form. Lacking experimental data both for the magnitude and
origin of an inflow boundary flux, we deal with the inflow boundary in an explicit
way that resembles an approximation of a periodic boundary condition. At a timestep $n$,
when solving the linear system for the concentration $c^{n+1}$ at the next timestep $n+1$,
we set the inflow flux $h_{\mathrm{in}}^{n+1} = h_{\mathrm{out}}^{n}$, where the
latter is the outflow flux
\begin{equation*}
    h_{\mathrm{out}}^{n} = \JJ^n\cdot\nn = (c^n\uu^n - D\nabla c^n)\cdot\nn
\end{equation*}
at the previous timestep. In the limit that the timestep $\Delta t\to 0$, we would be imposing
\begin{equation*}
    \intGin{(\JJ\cdot\nn)} = \intGout{(\JJ\cdot\nn)}.
\end{equation*}
This implies conservation of solute concentration at each timestep,
which is the desired behavior
as we endeavor to model the ventricles as a closed system.
Note that the total concentration would still increase in time,
because we impose the concentration profile in the photoconversion
region $\Omega_{\mathrm{pc}}$ (region of interest 1).

Using the introduced approaches of handling the in- and outflow boundaries,
the resulting abstract weak formulation for the cardiac-induced/no-cilia model is:
find $c\in W$ such that
\begin{equation}
    a(c, \phi) + b(c, \phi) = L(\phi), \quad\forall\phi\in W, \label{eq:adv_diff_weak_formB}
\end{equation}
with the definitions
\begin{align}
    a(c, \phi) &= \intO{\pdifft{c}\phi} - \intO{\JJ\cdot\nabla\phi},
    \label{eq:adv_diff_bilinear_form}\\
    b(c, \phi) &= \intGout{c(\uu\cdot\nn)\phi}, \label{eq:adv_diff_skew_form} \\
    L(\phi) &= \intGin{h_{\mathrm{in}}\phi}.\label{eq:adv_diff_linear_form}
\end{align}

\textbf{Baseline model.} Since both free-slip and cilia tangential
traction boundaries have the same no-flux boundary condition
in the advection-diffusion equation, the weak formulation of
the advection-diffusion equation for the baseline model is
identical to that of the cardiac-induced/no-cilia model,
defined by~\Cref{eq:adv_diff_weak_formB,eq:adv_diff_skew_form,eq:adv_diff_linear_form}.

\subsection{Discretization of the advection-diffusion equation}
We outline the discretization of the advection-diffusion weak formulation
for the baseline model, defined
by~\Cref{eq:adv_diff_weak_formB,eq:adv_diff_skew_form,eq:adv_diff_linear_form}.
The weak formulation is discretized with a symmetric interior-penalty method using
discontinuous Galerkin finite elements~\cite{Arnold1982AnElements}.
Exchanging the variables in~\Cref{eq:adv_diff_weak_formB} with their discrete
counterparts yields the discrete weak form: find $c_h\in W_h$ such that
\begin{equation*}
    a(c_h, \phi_h) + b(c_h, \phi_h) = L(\phi_h), \quad\forall\phi_h\in W_h.
\end{equation*}
The space $W_h$ is constructed with finite elements with
discontinuous Lagrange polynomial basis functions:
\begin{equation}
    W_h = \Big\{w_h : w_h\in L^2;  w_h\vert_K\in P_k(K); \forall K\in\Omega_h\Big\}.
\end{equation}
These elements allow for discontinuities in the discrete solution across element edges.
To ensure a stable discretization with these elements, we add the stabilization
\begin{equation}
    \begin{aligned}
        a_{\mathrm{stab}}(c_h, \phi_h)
        = &\sum_{f\in \mathcal{F}}\int_f{D\frac{\alpha}{\avg{h}}\jump{c_h}\cdot\jump{\phi_h}}\,\mathrm ds
        +\sum_{f\in \mathcal{F}}\int_f{\left({c_h \hat{u}_h}|^+ - {c_h \hat{u}_h}|^-\right)\cdot\jump{\phi_h}}\,\mathrm ds  \\
         -&\sum_{f\in \mathcal{F}}\int_f{D\avg{\nabla\phi_h}\cdot\jump{c_h}}\,\mathrm ds
         -\sum_{f\in \mathcal{F}}\int_f{D\avg{\nabla c_h}\cdot\jump{\phi_h}}\,\mathrm ds,
    \end{aligned}\label{eq:SIP_stabilization}
\end{equation}
to the discrete weak formulation, where
\begin{equation*}
    \hat{u}_h = \frac{1}{2}\left(\uu_h\cdot\nn + \vert\uu_h\cdot\nn\vert\right)
\end{equation*}
is used to upwind the velocity $\uu_h$ with respect to a
facet. The stabilization parameter $\alpha>0$ must be chosen
large enough to achieve stability. In~\Cref{eq:SIP_stabilization}, the first two terms
penalize jumps in the concentration flux across interior facets,
and are needed to preserve coercivity of the bilinear form.
The last two terms in $a_{\mathrm{stab}}$ are
added to preserve symmetry and consistency of the bilinear form, respectively.
The average and jump operators used
in~\Cref{eq:SIP_stabilization} are defined, respectively, as
\begin{equation*}
    \avg{c} = \frac{c^+ + c^-}{2} \qquad\mathrm{and}\qquad \jump{c} = c^+\nn^+ + c^-\nn^-.
\end{equation*}

By including the stabilization $a_{\mathrm{stab}}(c_h, \phi_h)$,
the final semi-discrete weak formulation of the advection-diffusion equation is:
find $c_h\in W_h$, such that
\begin{equation}
    a(c_h, \phi_h) + a_{\mathrm{stab}}(c_h, \phi_h) + b(c_h, \phi_h) = L(\phi_h),
    \quad\forall\phi_h\in W_h.
    \label{eq:discrete_weak_form_adv_diff}
\end{equation}
In time, we discretized~\Cref{eq:discrete_weak_form_adv_diff} with the
second-order backward difference scheme BDF2~\cite{Volker2016}. This scheme
approximates the time derivative as
\begin{equation*}
    \pdifft{c} \approx \frac{3c^{n+1} - 4c^n + c^{n-1}}{2\Delta t},
\end{equation*}
where $n$ denotes the timestep index. All of the other terms
in~\Cref{eq:discrete_weak_form_adv_diff} are evaluated at timestep $n+1$.
The method is implicit and of second order in time. We solved the fully
discrete weak formulation of the advection-diffusion equation
with the penalty parameter value $\alpha=50$.

%--------------VERIFICATION--------------%
\section{Verification}\label{sec:appB_verification}
\subsection{Method of Manufactured Solutions}\label{subsec:mms}
Let $\hat{\xx} = (\hat{x}, \hat{y}, \hat{z}) = \frac{1}{L}(x, y, z)$
be the spatial coordinates scaled by the length $L$, which is the length of
the computational domain $\Omega$ in the $x$-direction. Letting $\bm{\phi}$ be
the vector-valued function
\begin{equation*}
    \bm{\phi} = \left(\sin{(\pi(\hat{x} - \hat{y}))},
    \sin{(\pi(\hat{y} + \hat{z}))},
    \sin{(\pi(\hat{x} - \hat{z}))}\right)
\end{equation*}
and defining the velocity as 
\begin{equation}
    \uu_{\mathrm{mms}} = \nabla\times\bm{\phi},
    \label{eq:velocity_mms}
\end{equation}
we have by construction that the velocity is divergence-free,
owing to the identity $\nabla\cdot(\nabla\times\bm{F}) = 0$ for
an arbitrary vector field $\bm{F}$. The pressure was defined as
\begin{equation}
    p_{\mathrm{mms}} = \cos{(\pi(\hat{x} + \hat{y} + \hat{z}))}.
    \label{eq:pressure_mms}
\end{equation}
Using the Method of Manufactured Solutions~\cite{Roache2001CodeSolutions},
we construct a problem with the Stokes equations where~\Cref{eq:velocity_mms}
and~\Cref{eq:pressure_mms} are the solutions.
To this end, we consider the general form of the steady Stokes equations
\begin{subequations}
    \begin{align}
        -\nabla\cdot\bsig &= \bm{f} \qquad\mathrm{in \ }\Omega,\\
        \nabla\cdot\uu &= g \qquad\mathrm{\ in \ }\Omega,
    \end{align}
    \label{eq:stokes_mms}%
\end{subequations}%
with $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$ and
$\beps(\uu) = \frac{1}{2}\left(\nabla\uu + (\nabla\uu)^T\right)$, and the boundary condition
\begin{alignat}{2}
    \uu &= \uu_{\mathrm{mms}} \qquad&&\mathrm{on \ }\Gamma,
\end{alignat}
where $\Gamma=\partial\Omega$ is the boundary of the domain.
Since the velocity field defined by~\Cref{eq:velocity_mms} is divergence-free by construction,
$g = 0$. Setting the external force $\bm{f}=-\nabla\cdot\bsig(\uu_{\mathrm{mms}}, p_{\mathrm{mms}})$
using~\Cref{eq:velocity_mms} and~\Cref{eq:pressure_mms},
the solutions $\uu$ and $p$ to~\Cref{eq:stokes_mms} will be the ones
defined by~\Cref{eq:velocity_mms} and~\Cref{eq:pressure_mms},
as long as we enforce equality in the mean pressures, i.e. $\intO{(p_h - p_{\mathrm{mms}})}=0$.
Convergence rates are then readily calculated by comparing the numerical approximations of
$\uu$ and $p$ with $\uu_{\mathrm{mms}}$ and $p_{\mathrm{mms}}$
on meshes of varying grid resolution.

\subsection{Norms and convergence order}\label{subsec:norms_and_eoc}
Errors of the velocity $\uu_h$ and pressure $p_h$ approximated with finite elements
were measured in $\mathbf{H}^1$ and $L^2$ error norms, respectively.
With the exact velocity field $\uu_{\mathrm{ex}}$, the $\mathbf{H}^1$
error semi-norm was calculated as
\begin{equation*}
    \Hnorm{\uu_h - \uu_{\mathrm{ex}}}^2
    = \intO{\nabla(\uu_h - \uu_{\mathrm{ex}}) : \nabla(\uu_h - \uu_{\mathrm{ex}})}.
\end{equation*}
The $L^2$ error norms for the pressure were calculated as
\begin{equation*}
    \normltwo{p_h - p_{\mathrm{ex}}}^2 = \intO{(p_h - p_{\mathrm{ex}})^2}.
\end{equation*}
The approximated functions were interpolated onto elements in a space of piecewise-continuous
Lagrange polynomials of three polynomial degrees higher than the elements used to approximate
the functions, meaning that polynomial degrees 4 and 3 were used for the velocity and pressure
errors, respectively. Numerical integration was performed with a Gauss quadrature rule,
with the quadrature rule degree equal to the polynomial degree of the integrated
finite element functions.

Convergence rates reported in~\Cref{subsec:numerical_verification} were
approximated in the following way. Define the error $e_k$ for a given mesh
resolution $h_k$. Based on the assumption that the error scales as
$e_k \approx C{h_k}^l$, where $C$ is a constant independent of $h_k$,
the convergence order $l$ was approximated as
\begin{equation}
    l = \frac{\log{(e_k/e_{k+1}})}{\log{(h_k/h_{k+1}})}.
    \label{eq:convergence_order_estimate}
\end{equation}
Consider a finite element function 
\begin{equation*}
    \phi = \phi_i\psi_i(\xx)
\end{equation*}
with basis functions $\psi_i$, where $i$ are the degrees of freedom and Einstein
summation is employed.
The infinity norm of $\phi$ was approximated from the vector
of coefficients $\left\{\phi_i\right\}$, which represent point evaluations of $\phi$ as
\begin{equation*}
    \normlinf{\phi} = \max_i{\vert \phi_i\vert}.
\end{equation*}

\subsection{Numerical verification of model implementation}\label{subsec:numerical_verification}
A cylinder geometry meshed with varying degrees of refinement was used to
perform convergence analysis of the numerical methods used to solve the Stokes equations.
The cylinder had unit length and a diameter of one half length unit.
For the analysis, we used tangential traction boundary conditions on all
of the cylinder surface. Using the Method of Manufactured Solutions~\cite{Roache2001CodeSolutions}
with the velocity and pressure expressions defined in~\Cref{subsec:mms}
for $\uu_{\mathrm{ex}}$ and $p_{\mathrm{ex}}$, we calculated the error of 
the velocity approximation $\uu_h$
in the $\mathbf{L}^2$ norm and $\mathbf{H}^1$ semi-norm,
and the error of the pressure approximation $p_h$ in the $L^2$ norm (\Cref{tab:cylinder_error_rates}).
Based on the error calculations,
we calculated the experimental order of convergence,
as defined in \Cref{subsec:norms_and_eoc}.
We observe convergence in all of the reported error norms. Furthermore,
we report the error in calculation of the maximum velocities, and observe
convergence in this error as well. Finally, we calculated the $L^2$ norm of
the divergence $\nabla\cdot\uu_h$ to verify the property of mass conservation.
\begin{table}[!htbp]
    \footnotesize
    \centering
    \caption{Error norms for the convergence study
    performed with a cylinder mesh with
    the cilia-driven/no-cardiac flow model using $\mu=1$.
    For all error norms, convergence rates calculated
    with~\Cref{eq:convergence_order_estimate} are given
    in parentheses. The coarsest mesh had a mesh size $h_0$,
    and for every mesh refinement the mesh size $h$ was halved.}\label{tab:cylinder_error_rates}
    \begin{tabular}{cc|ccccc}
        \toprule
        $h/h_0$ & \# dofs &
        $\normltwovec{\uu_h - \uu_{\mathrm{ex}}}$
        & $\Hnorm{\uu_h - \uu_{\mathrm{ex}}}$
        & $\normltwo{p_h - p_{\mathrm{ex}}}$
        & $\vert\normlinf{u_h} - \normlinf{u_{\mathrm{ex}}}\vert$
        & $\normltwo{\nabla\cdot\uu_h}$\\ 
        \midrule 
        1    & 8652    & \num{2.3e-02}  (--)  & \num{4.3e-01}   (--)  & \num{5.6e-01}  (--)  & \num{2.4e-02}  (--)  & \num{1.4e-07} \\  
        1/2  & 41190   & \num{9.3e-03} (1.30) & \num{2.8e-01}  (0.61) & \num{3.9e-01} (0.53) & \num{5.0e-03} (2.25) & \num{2.8e-08} \\  
        1/4  & 255016  & \num{3.5e-03} (1.42) & \num{1.4e-01}  (1.00) & \num{2.1e-01} (0.85) & \num{1.4e-03} (1.83) & \num{7.3e-11} \\     
        1/8  & 1855337 & \num{1.3e-03} (1.46) & \num{6.9e-02}  (1.03) & \num{1.1e-01} (0.92) & \num{6.7e-04} (1.08) & \num{1.5e-10} \\   
        \bottomrule
    \end{tabular}
\end{table}

To verify the model implementation with the zebrafish brain ventricles mesh,
we compared velocity norms for varying degrees of refinement of the brain ventricles
mesh~(\Cref{tab:ventricles_norms_BDM}).
We used the original mesh obtained by meshing the surface geometry using
fTetWild~\cite{Hu2020FastWild}; the region around the photoconversion region is not
refined on the mesh used for the convergence study.
In the numerical experiments, we used the cilia
traction vector $\btau = \tau\lambda(\xx) P_{\nn}(\rr)$
defined in~\Cref{tab:r_vector_and_lambda_function}
and $\gamma=10$ for the BDM penalty parameter.

For all meshes, we report the maximum discrete velocity magnitude $\normlinf{u_h}$.
We also report $\normlinf{u_h}$ scaled by the total force $F=\intGc{\sqrt{\btau\cdot \btau}}$
applied on the cilia boundary $\Gc$. We chose this measure using $F$ as a scale because the regions
of the mesh marked as cilia-populated change slightly with every mesh refinement,
owing to the unstructured organization of the mesh cells.
These changes in the surface area over which the tangential traction is applied
result in changes to $\normlinf{u_h}$,
since the maximum velocity magnitude is a result of the total amount of force
$F$ applied on the cilia boundary.

To account for the fact that the volume of the domain $|\Omega|=\intO{}$ also changes
slightly with mesh refinement, we report the $\mathbf{L}^2$ norm of $\uu_h$ scaled by $|\Omega|$.
Finally, we provide verification of mass conservation by reporting the $L^2$ norms of
the divergence $\nabla\cdot\uu_h$. We remark that the units used in the model
implementation was millimeters, which is why~\Cref{tab:ventricles_norms_BDM}
reports values in units of millimeters, instead of micrometers which has been used elsewhere in the paper. 
\begin{table}[!htbp]
    \centering
    \small
    \caption{Velocity norms under mesh refinement
    of the brain ventricles mesh using the cilia-driven/no-cardiac flow model.}
    \label{tab:ventricles_norms_BDM}
    \begin{tabular}{cc|cccc}
        \toprule
        \# cells & \# dofs &
        $\normlinf{u_h}$ $\left[\frac{\mathrm{mm}}{\mathrm{s}}\right]$
        & $\frac{1}{F}\normlinf{u_h}$ $\left[\frac{1}{\mathrm{Pa\cdot mm\cdot s}}\right]$
        & $\frac{1}{\vert\Omega\vert}\normltwovec{\uu_h}$
        $\left[\frac{\mathrm{mm}}{\mathrm{s}}\right]$
        & $\normltwo{\nabla\cdot\uu_h}$ $\left[\frac{\mathrm{mm^3}}{\mathrm{s^2}}\right]$\\
        \midrule 
        132134  & 943820  & 0.026 & \num{3.8e3} & \num{2.2e-03} & \num{1.9e-16}\\
        180901  & 1289881 & 0.027 & \num{3.7e3} & \num{2.2e-03} & \num{3.0e-16}\\
        262598  & 1870703 & 0.027 & \num{3.7e3} & \num{2.3e-03} & \num{6.7e-16}\\
        395339  & 2811668 & 0.028 & \num{3.6e3} & \num{2.2e-03} & \num{2.0e-14}\\
        627439  & 4448761 & 0.027 & \num{3.5e3} & \num{2.2e-03} & \num{4.9e-14}\\
        \bottomrule
    \end{tabular}
\end{table}

\subsection{Discussion of the numerical approximation and convergence rates}
A priori error estimates for the
Brezzi-Douglas-Marini and discontinuous Galerkin discretization scheme
$\mathrm{BDM}_k$-$\mathrm{DG}_{k-1}$ are naturally 
derived in a mesh-dependent broken $H^k$-norm for $\uu$ and the $L^2$-norm for the pressure.
The respective error norms are expected to decay with order $k$~\cite{Cockburn2005AEquations}.
For the sake of simplicity,~\Cref{tab:cylinder_error_rates}
reports the velocity errors in the standard $H^k$ norm. Using polynomial degree  
$k=1$, we observe linear convergence in the velocity in the $\mathbf{H}^1$ error norm
and the pressure in the $L^2$ error norm on the cylinder mesh. For the ventricles meshes,
we observe consistent values of the reported quantities, verifying the model
implementation.

The $\mathrm{BDM}_1$-$\mathrm{DG}_0$ scheme that we employ for the Stokes equations
is computationally more expensive than a more conventional approach based on
e.g. Taylor-Hood discretization~\cite{Stenberg1990ErrorProblem}
($P_2-P_1$ elements) or lower-order enriched
elements~\cite{Brezzi2011MixedMethods},
such as the mini-element ($P_1-P_1$ with a velocity stabilization).
We opted for the BDM-DG scheme because of its exactly-mass-conserving property,
since numerical experiments proved that a divergence-free velocity field was
fundamental to attain a stable discretization of the advection-diffusion equation
in advection-dominated regions. Preliminarily, we attempted
to discretize the Stokes equations with 
Taylor-Hood elements, but with the resulting numerical approximation
of the velocity $\uu_h$, we were unable to achieve
stability of the advection-diffusion discretization.
Numerical experiments identified non-zero divergence of $\uu_h$
as the root cause of these stability issues.
On the standard computational mesh, the $L^2$ norm of $\nabla\cdot\uu_h$ was
in the order of $10^{-6}$, while the
maximum value of the divergence of the velocity was of order unity.
On a refined mesh with roughly six times as many
computational cells as the standard mesh, the numerical values of
$\normltwo{\nabla\cdot\uu_h}$ and $\normlinf{\nabla\cdot\uu_h}$
remained practically unchanged. This is a result
of the fact that incompressibility is only achieved in a weak sense when
employing Taylor-Hood elements. The advective part of the total flux
is
\begin{equation*}
    \nabla\cdot(c\uu) = c(\nabla\cdot\uu) + \uu\cdot\nabla c.
\end{equation*}
In the case that $\nabla\cdot\uu\neq 0$, a source (or sink) term for the concentration
is introduced. This may introduce errors in the numerical approximation of $c$
and lead to instabilities and spurious oscillations, especially in the case of
advection-dominated transport
\cite{gresho1998incompressible, cesmelioglu2022compatible, johnson2009numerical}.
The $\mathrm{BDM}_1$-$\mathrm{DG}_0$ scheme avoids these issues,
since $\nabla\cdot\uu_h=0$ is satisfied exactly as a result of the element definitions.

In addition to its favorable mass-conserving property, the BDM-DG scheme
lets us impose the impermeability condition strongly without any tailored approach,
since this boundary condition appears naturally in the weak form of the
Stokes equations~(\Cref{subsec:appendixA1}) and the BDM elements control
the normal component of the velocity. If instead (for example) Taylor-Hood
elements were to be used, the impermeability boundary condition must
be handled with for example a Nitsche method~\cite{Nitsche1971UberSind}
or with a Lagrange multiplier~\cite{Babuska1973TheMultipliers,
Bertoluzza2017BoundaryHemodynamics}.

\section{Quantification of velocities in particle tracking}\label{sec:appC_particle_transport}
Particle tracking data was analyzed~\cite{Olstad2019CiliaryDevelopment}.
For four zebrafish, five recordings with a few particles for each fish were carried out.
The particles tracked were in the duct between the rhombencephalic ventricle and the
diencephalic ventricle. Frequency of image acquisition was 20.22 Hz.
The collected data constituted $x$ and $y$ positions of the particles at every time
instant of image acquisition. The velocities $u_x$ and $u_y$ were determined by
calculating the change in positions $\Delta x$ and $\Delta y$ relative to the change
in time $\Delta t$ between two subsequent images, $\Delta t$ being equal to the
frequency of acquisition.
Mean velocities $\overline{u}_x$~(\Crefpanels{fig:particle_tracking_data}{a}) and
$\overline{u}_y$~(\Crefpanels{fig:particle_tracking_data}{b}),
which serve as a quantification of particle drift, were calculated
by taking the difference between the last and first positions of a particle.

Particle movement varied over the different measurements. The mean of $\overline{u}_x$ was
0.26 \textmu m/s and the mean of $\overline{u}_y$ was 0.08 \textmu m/s.
The maximum mean velocities
in absolute values were 10.2 \textmu m/s and 2.4 \textmu m/s for
$\overline{u}_x$ and $\overline{u}_y$,
respectively. We used the cardiac-induced/no-cilia flow model to tune the
pressure boundary condition
amplitude $A$. The choice of $A=1.5$ mPa was used since: (i) it yields a
maximum velocity magnitude
of 4.8 \textmu m/s, which is close to the maximum observed in the
particle tracking data when
rejecting the outlier at around 10 \textmu m/s; (ii) it yields a time-averaged mean velocity
of 0.89 \textmu m/s, which is close to the mean observed in the particle tracking data.
\begin{figure}[H]
    \begin{minipage}[t]{0.49\textwidth}
        \includegraphics[width=\textwidth]{graphics/mean_velocities_x.png}
    \end{minipage}
    \begin{minipage}[t]{0.49\textwidth}
        \includegraphics[width=\textwidth]{graphics/mean_velocities_y.png}
    \end{minipage} 
    \caption{Particle tracking data. Mean velocities in
    the $x$ (\textbf{a.}) and $y$ (\textbf{b.}) directions of the particles tracked.
    The dashed horizontal line designates a value of zero.}\label{fig:particle_tracking_data}
\end{figure}

\end{document}